### 2016-09-26 journal-draft/bare_jrnl_comsoc.tex
modified the centralised section

\documentclass[10pt,journal,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[journal,comsoc]{../sty/IEEEtran}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% copied from thesis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Packages für Formeln %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[colorlinks,
            linkcolor=red,
            anchorcolor=blue,
            citecolor=green
            ]{hyperref}
\usepackage{amsmath}
\usepackage{amsthm}
%\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{booktabs}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

%% Packages
%\usepackage{tikz}
%\usetikzlibrary{shapes,arrows}
%\usetikzlibrary{positioning}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{kantlipsum} %<- For dummy text
\usepackage{mwe} %<- For dummy images
\usepackage{multirow}
\usepackage{multicol}
%\usepackage{colortbl}
%\usepackage{todonotes}	
\newcounter{todocounter}
\newcommand{\todonum}[2][]
{\stepcounter{todocounter}\todo[inline,backgroundcolor=blue!10!white, #1]{\thetodocounter: #2}}

%% Zeilenabstand %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{setspace}
%\singlespacing        %% 1-zeilig (Standard)
%\onehalfspacing       %% 1,5-zeilig
%\doublespacing        %% 2-zeilig

%% Andere Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{a4wide} %%Kleinere Seitenränder = mehr Text pro Zeile.
\usepackage{fancyhdr} %%Fancy Kopf- und Fußzeilen
%\usepackage{longtable} %%Für Tabellen, die eine Seite überschreiten
\usepackage{lscape}
\usepackage{rotating} 
%\usepackage[htt]{hyphenat} %Trennung von Typewriter-Schriften
%\usepackage{listings}
%\usepackage{pstricks-add} --> This package generates problems with booktabs (toprule, etc.)
\usepackage[autostyle]{csquotes}
\usepackage{amsmath,bm}
\usepackage{array}
% Tabellen mit Center und left
\usepackage{tabularx,colortbl} % colored table background
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
% Table spacings
\newcommand\T{\rule{0pt}{2.5ex}\rule[-1.0ex]{0pt}{0pt}}
\newcommand\B{\rule[-1.0ex]{0pt}{0pt}}

\definecolor{slightgray}{gray}{.90} 
\usepackage{rotating}
\usepackage{hhline}
\usepackage{float}
\usepackage{caption}% http://ctan.org/pkg/caption
\captionsetup[table]{format=plain,labelformat=simple,labelsep=period}%

\newtheoremstyle{mytheoremstyle} % name
        {\topsep}                    % Space above
        {\topsep}                    % Space below
        {\itshape\fontfamily{ptm}\selectfont}                   % Body font, ptm is times new roma
        {}                           % Indent amount
        {\fontfamily{ptm}\selectfont\scshape\color{black}\bfseries}                   % Theorem head font
        {:}                          % Punctuation after theorem head
        {.5em}                       % Space after theorem head
        {}  % Theorem head spec (can be left empty, meaning ‘normal’)

\theoremstyle{mytheoremstyle}
\newtheorem{theorem}{Theorem}[section]

\theoremstyle{mytheoremstyle}
\newtheorem{corollary}{Corollary}[section]

\theoremstyle{mytheoremstyle}
\newtheorem{lemma}{Lemma}[section]

%\theoremstyle{mytheoremstyle}
%\let\definition\relax
\newtheorem{mydef}{Definition}%[section]

%%%%%%
%%%%%% Proof
    \makeatletter
    \renewenvironment{proof}[1][\proofname]{%
      \par\pushQED{\qed}\fontfamily{ptm}\selectfont%
      \topsep6\p@\@plus6\p@\relax
      \trivlist\item[\hskip\labelsep\bfseries#1\@addpunct{.}]%
      \ignorespaces
    }{%
      \popQED\endtrivlist\@endpefalse
    }
    \makeatother 
%%%%%%%%%%%%%%%%%

\newcommand{\eg}{e.g., }
\newcommand{\ie}{i.e., }
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
%\usepackage{algorithmic}
\usepackage{times}
%\usepackage{mathptmx}
\usepackage{mathtools}
\usepackage[draft,nomargin,marginclue,footnote,silent]{fixme}
\setcounter{tocdepth}{2} %table of contents

\makeatother
\usepackage{comment}
\newcommand{\Tau}{\mathrm{T}}
\usepackage[colorinlistoftodos]{todonotes}
\DeclareMathOperator*{\Max}{Max}
\DeclareMathOperator*{\Min}{Min}

\newcommand{\bigO}{\ensuremath{\mathcal{O}}}% big-O notation/symbol
\usepackage{subfigure}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{lipsum}
\setlist[itemize]{leftmargin=*}
%%%%%%%%%%%%%%%%%%%%%%%%%% copied from thesis
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[T1]{fontenc}% optional T1 font encoding

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
% Do NOT use the amsbsy package under comsoc mode as that feature is
% already built into the Times Math font (newtxmath, mathtime, etc.).
% 
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% Select a Times math font under comsoc mode or else one will automatically
% be selected for you at the document start. This is required as Communications
% Society journals use a Times, not Computer Modern, math font.
%\usepackage[cmintegrals]{newtxmath}
\usepackage{txfonts}
% The freely available newtxmath package was written by Michael Sharpe and
% provides a feature rich Times math font. The cmintegrals option, which is
% the default under IEEEtran, is needed to get the correct style integral
% symbols used in Communications Society journals. Version 1.451, July 28,
% 2015 or later is recommended. Also, do *not* load the newtxtext.sty package
% as doing so would alter the main text font.
% http://www.ctan.org/pkg/newtx
%
% Alternatively, you can use the MathTime commercial fonts if you have them
% installed on your system:
%\usepackage{mtpro2}
%\usepackage{mt11p}
%\usepackage{mathtime}

%\usepackage{bm}
% The bm.sty package was written by David Carlisle and Frank Mittelbach.
% This package provides a \bm{} to produce bold math symbols.
% http://www.ctan.org/pkg/bm

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Versatile Robust Clustering of Ad Hoc Cognitive Radio Network}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{Di~Li,~\IEEEmembership{Member,~IEEE,}
        Erwin~Fang,
        and James~Gross,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
\thanks{D. Li was with RWTH Aachen university, Germany.}% <-this % stops a space
\thanks{Erwin. Fang is with ETH, Switzerland.}% <-this % stops a space
\thanks{J. Gross is with KTH Royal Institute of Technology, Sweden.}% <-this % stops a space
\thanks{Manuscript received xxxx xx, 20xx; revised xxxx xx, 20xx.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of xxxxx,~Vol.~xx, No.~x, xxxx~20xx}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Communications Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2015 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
Cluster structure in cognitive radio networks facilitates cooperative spectrum sensing, routing and other functionalities.
The unlicensed channels, which are available for every member of a group of cognitive radio users, consolidate the group into a cluster, and the availability of unlicensed channels decides the robustness of that cluster against the licensed users' influence.
%Thus there should be more unlicensed channels in the clusters.
This paper analyses the problem that how to form robust clusters in cognitive radio network, so that more cognitive radio users can get benefits from cluster structure even when the primary users' operation are intense.
%In the process of forming clusters, every secondary user decides with whom to form a cluster, or which cluster to join.
We provide a formal description of robust clustering problem, prove it to be NP-hard and propose a centralized solution, besides, a distributed solution is proposed to suit the dynamics in the ad hoc cognitive radio network.
Congestion game model is adopted to analyse the process of cluster formation, which not only contributes designing the distributed clustering scheme directly, but also provides the guarantee of convergence into Nash Equilibrium and convergence speed.
Our proposed clustering solution is versatile to fulfill some other requirements such as faster convergence and cluster size control.
The proposed distributed clustering scheme outperforms the related work in terms of cluster robustness, convergence speed and overhead.
%Besides, we prove the clustering problem is NP-hard, and also propose the centralized solution.
The extensive simulation supports our claims.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
cognitive radio, robust cluster, game theory, congestion game, distributed, centralized, cluster size control.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\graphicspath{
{../figures/04_clutering/}
}

\section{Introduction}
\label{intro}
%CR concept
\IEEEPARstart{C}{ognitive} radio (CR) is a promising technology to solve the spectrum scarcity problem~\cite{Mitola}.
Licensed users access the spectrum allocated to them whenever there is information to be transmitted.
In contrast, as one way, unlicensed users can access the spectrum via opportunistic spectrum access, \ie they access the licensed spectrum only after validating the channel is unoccupied by licensed users, where spectrum sensing~\cite{sensing_survey_2009} plays an important role in this process.
%
In this hierarchical spectrum access model~\cite{zhao_survey_DSA_2007}, the licensed users are also called primary users (PU), while the unlicensed users are referred to as secondary users and constitute a so called cognitive radio network (CRN).
%
%cluster
Regarding the operation of CRN, efficient spectrum sensing is identified to be critical for a smooth operation of a cognitive radio network~\cite{Sahai_FundamentalDesignTradeoffs2006}.
This can be achieved by cooperative spectrum sensing of multiple secondary users, which has been shown to cope effectively with noise uncertainty and channel fading, thus remarkably improving the sensing accuracy~\cite{coorperativeSensing_Akyildiz11}.
Collaborative sensing relies on the consensus of CR users\footnote{The terms user and node appear interchangeably in this paper. In particular, user is adopted when its networking or cognitive ability are discussed or stressed, while we refer node typicallly in the context of the topology.} within a certain area, in this regard, clustering is regarded as an effective method to realize cooperative spectrum sensing~\cite{Sun07_clustering_spectrum_secsing, Zhao07}.
% ... decreases considerably the negative probability due to the random channel effects like fading and shadowing.
Clustering is a process of grouping certain users in a proximity into a collective.
Clustering is also efficient to coordinate the channel switch operation when primary users are detected by at least one CR node residing in the cluster.
The cluster head (CH) can enable all the CR devices within the same cluster to stop payload transmission swiftly on the operating channel and to vacate the channel ~\cite{willkomm08}.
%
%As to the coexistence of CR users, they can be notified by cluster head (CH) or other cluster members about the possible collision, the possibility for them to interfere neighbouring clusters is reduced~\cite{centralizedSharing80222}. 
In addition to the collaborate sensing advantage, the use of clusters is beneficial as it reduces the interference between cognitive clusters~\cite{centralizedSharing80222}.
Clustering algorithm has also been proposed to support routing in cognitive radio networks ~\cite{Abbasi_survey_07}.

%%[crn for clustering] 
%%due to attenuation of signal propagation, primary users can only be detected by CR users when they locate closely to CR users.
%In cognitive radio networks, secondary users which locate closely with each other are possibly affected by the same group of primary users, so that the availability of licensed spectrum is similar to them, \ie certain channels are available on each of them.
%The similarity of available spectrum on a group of neighbouring CR nodes, along with the benefit of collaborative decision among multiple nodes, leads to clustering as an effective approach for many applications.
%[robustness issue for clustering]

Forming clusters is conducted in the very beginning or periodically according to the dynamics of the CRN.
To form a cluster which is composed by the CR users, there should be channels available for communication within the cluster.
Usually there are multiple unlicensed channels available for all the CR nodes in the cluster, which are referred to in the following of this paper as \textit{common licensed channels} (or common channels for short, which is abbreviated as CC).
Out of the available CCs, there are always one or multiple channels which are used for the payload communication.
%The communication within a cluster is conducted in the spectrum which is available for every member in that cluster.
When one or several cluster members can not use one certain CC for example because primary user activity has been detected on that channel, it will be excluded from the set of CCs.
If the channel is being used for payload communication, all the cluster members will switch to another channel in the set of available CCs to achieve seamless transmission.
As long as there is a CC available, the corresponding cluster can continue to exchange payload data.
In the context of CRN, as the activity of primary users is unknown to the secondary users, the availability of CCs of the formed clusters is totally dependent on primary users' activity.
Thus, the clusters with more CCs are in general better as this anticipates a longer time span that payload can be exchanged.
In this paper the robustness of a cluster against primary users is synonymous with the affluence of CCs which are possessed by that cluster. 
%In other words, the availability of CCs of the formed clusters is passive and can not be guaranteed.
%Thus, the number of CCs in one formed cluster indicates the robustness of it when facing the unpredictable influence from primary users.
%In CRN, one cluster survives the influence of primary users when at least one CC is available in that cluster.
%As the primary users' operation is assumed to be unknown to the CR users, a cluster formed with more CCs will survive with higher probability.

%It is easy to see that forming clusters with different neighbours results in different number of CCs in the clusters.
%As a result, how to form the clusters plays an important role on the robustness of clusters in CRN.
%Furthermore, the clustering algorithm also determines the connectivity between several clusters which ultimately determines the robustness of the entire network on the respect of connectivity. 
In the formation process, there is a trade-off between CCs and the collaborative sensing performance, as less members in a cluster yield in general a higher number of CCs while this reduces the sensing accuracy~\cite{Consensus_based_clustering12}. 
In addition, there are other reasons that motivate clusters to have more members.
For example, instead of less members, the proper cluster size results in smaller power consumption~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
%
%To solely pursue cluster robustness against the primary users' activity, \ie to achieve more common channels within clusters, the ultimately best clustering strategy is ironically that every node constitutes one cluster (called as \textit{singleton cluster} in this paper).
%%In that case, the common channels within cluster are the available channels available at that node's place.
%Apparently this contradicts our motivation of proposing cluster in cognitive radio network to enable cooperative decision making.
Thus the cluster robustness discussed in terms of number of CCs carries little meaning when the sizes of formed clusters are not given consideration.

There has been a lot of research done for clustering in wireless networks.
%Prior to the emergence of cognitive radio, clustering has been proposed as an important method to manage network ad-hoc networks~\cite{Kawadia03,Lin97adaptiveclustering,Basagni99}, wireless mesh networks\cite{clustering_mesh_globecom2010} and sensor networks~\cite{Abbasi_survey_07}. 
In ad-hoc and mesh networks, the major goal of clustering is to maximize connectivity or to improve the performance of routes~\cite{Kawadia03, clustering_mesh_globecom2010}.
The emphasis of clustering in sensor networks is on network lifetime and coverage~\cite{Abbasi_survey_07}.
%Few works also take the overhead of clustering into account in ad-hoc network~\cite{clusterRoutingOverhead02infocom, clusterRoutingOverhead_wcnc04}.
Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
The clustering schemes proposed in~\cite{Zhao07, Chen07,Affinity_clustering_09icccn} form clusters which meet the basic condition, having CC in every cluster.
Clustering scheme~\cite{Consensus_based_clustering12} improves spectrum sensing accuracy.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{clustering_globecom11, TWC2012_cooperative_communication} target on the QoS provisioning and energy efficiency.
\cite{cluster_EW10} forms clusters to coordinate the control channel usage within one cluster.
A event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
% optimize the sizes of the formed clusters to decrease power consumption.
%the channel available to the largest set of one-hop neighbours is selected as common channel which yields a partition of the CRN into clusters. 
%This approach minimizes the set of distinct frequency bands (and hence, the set of clusters) used as common channels within the CRN.
%However, bigger cluster sizes generally lead to less options within one cluster to switch to if the common channel is reclaimed by a primary node. 
%This method pays attention to the communication among the formed clusters, but cluster robustness is not considered.
%\cite{TWC2012_cooperative_communication} targets on the QoS poisoning and energy efficiency. 
%This approach first decides on the relay nodes which minimize transmission power consumption, then the chosen nodes become cluster heads and clusters are formed in a dynamic coalition process.
%This work emphasis on power efficiency and doesn't take into account the channel availability and the issue of robustness of the formed clusters.
%\cite{cluster_EW10} deploys cluster structure in order to implement common channel control, medium access  with multiple channel and channel allocation. 
%The node with the maximum number of common channels within its k-hop neighborhood is chosen as cluster head, but how to avoid one node appearing in multiple clusters is not given consideration.
%A event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
%The cluster heads are selected according to node degree, spectrum availability and the distance to the sink, and the clusters are no longer available after the event ends.
No one among the above mentioned schemes provides the robustness to the formed clusters against primary users.

%Forming robust clusters is considered in~\cite{LIU_TMC11_2, mansoor_15_cluster_robust}.

A clustering scheme (denoted by SOC) which is designed to generate robust clusters against primary users is proposed in~\cite{LIU_TMC11_2}.
SOC involves three phases of distributed executions.
In the first phase, every secondary user forms clusters with some one-hop neighbors, in the second and third phase, each secondary user seeks to either merge other clusters around it, or join one of them.
The metric adopted by every secondary user in the three phases is the product of the the number of CCs and cluster size (the number of secondary users in the cluster).
%To the best of our knowledge, SOC is the only work emphasizing on the robustness of clustering structure among the related works. 
But this robust clustering scheme has some drawbacks.
Although the adopted metric covers both cluster size and the number of CCs, the formation of certain clusters may be easily dominated by only one factor, e.g. a node which is able to use many channels will exclude its neighbor and form a cluster by itself.
%emphasis on improving the numbers of common channels within clusters, in order to strengthening robustness of clusters, but the perused metric is not examined or proved to be able to maintain cluster structure under vigorous intensity of primary users’ activities.
%The authors consider the balance between the number of idle common channels within cluster and cluster size and propose an algorithm that increases the number of common channels within clusters. 
%However, this work neglects the issue of connectivity between clusters. 
%One drawback of this scheme is, in order to increase the number of common channels within clusters, the scheme excludes certain CR nodes from the formed clusters, so that isolated nodes have to form clusters themselves. 
In addition, this scheme leads to the high variance of the cluster sizes, which is not desired in certain applications as discussed in~\cite{clustering_globecom11, cluster_EW10}.
%The first two propose a distributed scheme where the metric is the product of cluster size and the number of common channels.
\cite{mansoor_15_cluster_robust} presents a heuristic method to form clusters, although the authors claim robustness is one goal to achieve, the minimum number of clusters is pursued.
\cite{Li11_ROSS} proposes a distributed clustering scheme under the game theoretic framework. 
Compared with the clustering schemes introduced so far, the clusters are formed in shorter time and posses more CCs within and among clusters.
But this scheme doesn't consider the the factor of size in the algorithm design.

In this paper we will give a comprehensive analysis of the robust clustering problem and propose solutions. 
We stick to the motivation of forming robust clusters in CRN \ie let more CR users benefited from the cooperative decision making which is due to the clusters. 
We propose both centralized and distributed clustering schemes, which result in more CR users in the clusters composed with multiple CR users, not only when the clusters are just formed, but also afterwards when the primary users' activity changes.
Besides, both centralized and distributed schemes take the cluster size into consideration, and both can control the sizes of the formed clusters to certain extent.
In particular, the decentralized clustering approach ROSS (RObust Spectrum Sharing) is able to form clusters with desired sizes.
When compared with other distributed works, ROSS involves smaller signaling overhead and the generated clusters are significantly more robust against the primary users which appear after the clusters are formed.
%thesis
%and the generated clusters are more robust than the other clustering scheme which has claims on cluster robustness, \ie more secondary users residing in clusters against increasing influence from primary users.
%
%ROSS selects cluster head through coordination within its neighborhood, and then cluster membership is decided locally and its fast convergence is proved under game theoretic framework. 
%For our scheme we can prove convergence in cluster formation phase and resolve ambiguities with respect to cluster membership in a game-theoretic setting. 
We also propose the light weighted versions of ROSS, which involve less overheads and thus are more suitable for the scenario where fast deployment is desired.
Throughout this paper, we refer the clustering schemes on the basis of ROSS as \textit{variants of ROSS}, \ie the fast versions, or that with size control feature.

The rest of paper is organized as follows. 
We present the system model in Section~\ref{sec:model}. 
Then the formal description of robust clustering problem in CRN and the properties of the problem are given in Section~\ref{problem}.
The centralized and distributed solutions are introduced in Section~\ref{centralized_opt} and \ref{ross} respectively.
%The clustering problem is given through analysis and a centralized scheme is proposed in section~\ref{centralized_scheme}.
Extensive performance evaluation is presented in Section~\ref{performance}.
Finally, we conclude our work and point out the direction for future research in Section~\ref{conclusion}.

\section{System Model and Problem Formulation}
\label{sec:model}
%Let us consider a two dimensional area where primary and secondary users coexist together.
%They share the licensed channels according to the spectrum overlay model, where secondary transmitters are only allowed to transmit when the prifcan exchange its spectrum sensing resulmary users are detected as being idle, and they should vacate the working channel when the presence of primary user is detected.
%The set of primary users and secondary users are presented by $\mathcal{P}$ and $\mathcal{N}$ separately, there are $|\mathcal{P}| = P$ and $|\mathcal{N}| = N$.
%Cognitive radio ad hoc network is constituted by all the secondary users in $\mathcal{N}$.
%The collection of non-overlapping licensed frequency bands is denoted as $\mathcal{F}$ with $|\mathcal{F}| =F$, which is shared by the the primary and secondary users.
%We assume that primary users have a relatively low variation in activity (periods of activity and inactivity in the range of seconds or minutes).
%%
%Primary users access the allocated channels in $\mathcal{F}$ according to its need without sending any explicit notification to secondary users.

We consider a set of cognitive radio users $\mathcal{N}$ and a set of primary users distributed over a given area.
%We denote the set of these nodes as $V$ which is partitioned into sets of primary users $V_P$ and secondary users $V_N$, i.e., $V~=~V_P~\dot\cup~V_N$. 
%These users share a number of non-overlapping licensed channels according to the spectrum overlay model. 
The CR users use a set of licensed channels opportunistically, which is denoted as $\mathcal{K}$. 
The CR users are allowed to transmit on channel $k \in \mathcal{K}$ only if no primary user is detected on channel $k$. 
%The available channels sensed on secondary user $i$ is denoted by $V_i$ and there is $\vert V_i \vert \leq F$. % indicating the total number of 
%%Each secondary user notifies its neighbours of the sensing result about the available channels.
CR users conduct spectrum sensing independently and sequentially on all licensed channels.
%The sensing duration and frequency on one channel is a research topic~\cite{sensing_survey_2009}, and 
We assume that every node can detect the presence of an active primary user on each channel with certain accuracy. \footnote{The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
We adopt the unit disk model~\cite{unitDiskModel} for the transmission of both primary and CR users.
%Both primary users and CR users have fixed transmission ranges respectively, and all the channels are regarded to be identical in terms of signal propagation.
If a CR node locates within the transmission range of an active primary user, that CR node is not allowed to use the channel which is being used by that primary user.
%In our system model, we assume that a secondary node $n$ is able to detect for each channel $c$ whether any primary user utilizes $c$.
As the result of spectrum sensing, $K_i \subseteq \mathcal{K}$ denotes the set of licensed channels which can be used by $i$.
%
%One dedicated control channel is assumed to be available for all the secondary nodes to exchange control messages with neighbours in the process of cluster formation.
%Note that the assumption of dedicated control channel is to simplify the discussion so that we can focus on the kernel of this paper, the robustness of clusters.
%Actually, the control messages involved in the clustering process can be conveyed on available licensed channels through rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}. 
%Over the control channel, secondary users exchange their spectrum sensing results $V_{i}$ with one hop neighbours. 
%Secondary users have the same transmission range on both licensed and control channel, and the communication links are reciprocal.
%

We assume there is one dedicated control channel which is used to exchange signaling messages during clustering process.
This control channel could be one of the ISM bands or other reserved spectrum which is exclusively used for transmitting control messages.
\footnote{Actually, the control messages involved in the clustering process can also be transmitted on the available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighbouring nodes establish communication on the same channel.}
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ to all its one hop neighbors $\text{Nb}(i)$, where $\text{Nb}(i)$ is the neighbors of $i$, and is simply defined as the set of CR nodes located within the transmission range of $i$. 
%It is available for any secondary node $i$ to exchange control messages with any other node in its neighborhood $\text{Nb}(i)$ during the cluster formation phase.
In the following, we refer to the word \textit{channel} in general as licensed channel, unless we mention the control channel explicitly. 
%
%
If a secondary user $i$ is not in the transmission range of an active primary user $p$, $i$ can certainly not detect the presence of $p$.
As the transmission range of primary users is limited and secondary users have different locations, different secondary users may have different views of the spectrum availability, i.e., for any $i, j \in \mathcal{N}$, $K_i = K_{j}$ does not necessarily hold.	
%
%Neighbourhood establishment and maintenance with control channel and according to a neighborhood discovery protocol which is out of scope of our work.
%While primary users are assumed to be fixed, secondary users can be mobile.
%Validation refers to the process of ensuring that no primary transmission is actually taking place on the respective channel.
%All $N$ secondary users constitute an ad-hoc network in which data can be transmitted from one certain node to any other node should available c.hannels are available on the source node, destination node and other intermediate nodes.
%Communication on licensed channels is possible only when nodes $i, j$ are both located in each other's transmission range and both share a validated licensed channel.
%As the assumed $0/1$ state of connectivity is solely based on the Euclidean distance between secondary users,
%CRN can be represented by a graph $\mathcal{G}(I,\mathcal{E})$, where $\mathcal{E}=\lbrace(i,j,v) \vert i, j \in \mathcal{N} \wedge v\in V_i \wedge v\in V_j \rbrace$ is wireless link between any secondary node $i$ and its neighbour $j$ with licensed channel $v$.
%Due to relatively low primary user dynamics, time index is omitted here.
%%In order to perform clustering, CR users first need to establish their neighborhood.
%For secondary node $i$, its neighborhood $Nb_i$ consists of all the secondary users locating within its transmission range, regardless whether common licensed channels exist or not. 
%% and have at least one common channel with node $i$ each, i.e. $ j\in Nb_i \Rightarrow V_i\cap V_j\neq \emptyset$.
%%The clustering phase is initialized during which any control message is again conveyed by the control channel.
%In the rest of this paper, \textit{channel} only refers to the licensed common channel unless the dedicated control channel is particularly mentioned.
Then, a cognitive radio network can be represented as a graph $G = (\mathcal{N}, E)$, where $E \subseteq \mathcal{N} \times \mathcal{N}$ such that $\{i, j\} \in E$  if, and only if, there exists a channel $k \in \mathcal{K}$ with $k \in K_{i} \cap K_{i'}$. 
%Note that the channel availability is that when the clustering process is performed. 
%For the rest of this paper the word channel is referred to licensed channel, if the control channel is not explicitly mentioned.

%In this subsection, we present the definition of cluster in the context of CRNs, define the clustering problem and discuss its complexity.
%We give the description of cluster in the context of CRN, all the clusters discussed in this paper comply with it.
%A cluster $C$ is composed with one cluster head and cluster members, which satisfies the following conditions:
%\begin{itemize}
%\item Cluster head $H_C$ is able to communicate with any cluster member \textit{directly}, \ie for any cluster member $i\in C$, there is $i\in \texttt{Nb}_{H_C}$.
%\item There exists as least one common channel in the cluster, \ie $\cap_{i\in C} V_i \neq \emptyset$.
%\end{itemize}
A cluster $C$ is a set of secondary nodes, which consists a cluster head $h_C$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
In other terms, for any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
%
%A cluster head coordinates the activities of cluster members, \ie it notifies all the members to vacate a channel if a primary user's presence is detected by one cluster member. Further it also notifies the members to use a different common channel for payload communication. %(\textcolor{red}{(Not clear what the meaning of payload communication is. Why not omitting payload here?)} \textcolor{brown}{payload communication happens when the clusters are formed, and start to support network functionalities, e.g. routing. Thus I think it is necessary to leave it here.}
%
%
In this paper, sometimes we need to mention a cluster with the information of its cluster head \ie a cluster is denoted as $C(i)$ where $i$ is its cluster head.
Cluster size of $C(i)$ is written as $|C(i)|$.
$ K(C) = \cap_{i\in C} K_i$, $K(C)$ denotes the set of common licensed channels in cluster $C$.
Clustering is performed periodically, as secondary users are mobile and the channel availability on secondary users are changing as primary users change their operation state.%, or the most current clusters can not maintain due to lack of CCs when primary users' operation is intense.

The notations used in the system model and the following problem description are listed in the Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & collection of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h_C$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
$C_i$  & the $i$th potential cluster in a collection of clusters \\
\bottomrule
\end{tabular}
\end{table}

%We propose distributed scheme ROSS to form clusters to generate robust clusters, and in the same time clusters have preferred sizes, \ie fewer number of singleton clusters (the cluster which consist with only one CR node) compared with state of art.

%Besides, cluster size is also considered in to clustering solution.
%Size preference can be met after minor modifications on ROSS.

%The metric is summation of the number of channels available to be used for each node when they reside in a certain cluster, together with a cost for not following the desired cluster size, \ie when the desired cluster size is $\delta$ and the other cluster sizes are denoted as $\delta'$, the metric is, 
%\begin{equation}
%\label{metric}
%\sum_{i=1}^{N}(ICC_i)
%\end{equation}
%$N$ is the number of CR nodes in the network.

%Also, we refer to the common channels between neighbouring clusters by the term \textit{outward common channels} (OCC). We define the set of OCCs of cluster $C$ to be the set of available common channels between any member of $C$ and any other CR user of a neighbouring cluster:
%\begin{equation}
%\label{numocc}
%R_{C}=\bigcup_{j\in C, k\in N_j, k\notin C}(V_{j}\cap V_{k})
%\end{equation}

%In the process of clustering, when there is no restriction on cluster size, the metric will only be the summation of the number of channels for each node when they reside in a certain cluster.
%When we try to maximize the summation of ICCs, an obvious correlation between cluster size and number of ICCs is encountered, \ie when each node constitutes one cluster, the aforementioned metric will be maximized.
%The goal of the proposed clustering scheme is to let as more CR nodes as possible to form clusters, meanwhile the clusters have more common channels and preferred cluster size.

%\textcolor{red}{The notation is not well chosen. Does it make sense to have only one letter for a potential cluster? What happens if we have two clusters? Find some comments next to your proposed notation.}

\subsection{Robust Clustering Problem in CRN}
\label{problem}
%paper given by james is useful which provides a survey from the perspective of wsn!
As introduced in the section of introduction, in order to be robust against primary users' activity, the formed clusters should have more CCs to expect longer time expectancy.
Meanwhile, the sizes of the formed clusters should not diverge from the desired size too much, and there are two reasons.
First, the generation of the \textit{singleton clusters}, \ie the cluster which has only one CR node, should be avoided, because it contradicts the motivation of forming clusters thus the benefits brought in by the collective of the cluster members are compromised.
Second, Large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members is a substantial burden. 
Hence, the cluster size should fall in a desired range according to different application scenarios~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
Considering the above mentioned requirements on the formed clusters, we will present the definition of robust clustering problem as follows.
%The clustering scheme we propose in the paper aims to form clusters with preferred sizes, meanwhile maximize the average number of CCs per cluster.
%We have mentioned that the sizes of the formed clusters should be given consideration, which is import for the discussion on cluster robustness, \ie 
%When perusing the cluster robustness without specifying the requirement on cluster sizes, a clustering method will be inclined to form small clusters.
%In that case, although more CCs are obtained per cluster, the benefits brought in by the collective of the cluster members are compromised.
%Except for cooperative sensing, clusters need to conduct some other functionalities.

\begin{mydef}
\label{def_centralized_clustering}
\textit{Robust clustering problem in CRN.}

Given a cognitive radio network $\mathcal{N}$ where $|\mathcal{N}|=N$,  the secondary users are indexed from 1 to $N$.
A cluster $C$ is formed when first, it is composed with one or multiple secondary users and there is $|K(C)|>0$, second, any two cluster members can communicate either directly, or through other members which belong to the same cluster.
The collection of all the possible clusters in $\mathcal{N}$ is denoted as $\mathcal{S}$, there is $\mathcal{S}=\{C_1, C_2,\ldots,C_{|\mathcal{S}|}\}$ where $C_i$ is the $i$ th cluster in $\mathcal{S}$\footnote{The sequence of the clusters can be decided in any convenient way \ie the sequence of identifying them.}.
%the clustering solution will be composed by the elements of $\mathcal{S}$. 
As to $\mathcal{S}$, there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = N$.

To formulate the problem, we propose a new definition for the number of CCs in a cluster.
We use $f(C_i)$ to denote the number of CCs of a cluster $C_i\in \mathcal{S}$, which is as follows,
%and the number of common licensed channels is $f(C_i) = |K(C_i)|$ if $|C_i|>1$, and $f(C_i)=0$ when $|C_i|=1$.
$$
f(C_i) = \left\{ \begin{array}{rl}
|K(C_i)| 	&\mbox{$|C_i|>1$} \\
0 				&\mbox{$|C_i|=1$} \\
\end{array} \right.
$$

%The size of cluster $C$ is not bigger than a positive integer $k$.
%Then the sum of the number of CCs of the formed clusters is $\sum_{C\in \mathcal{S}'} f(C)$

Then the robust clustering problem is to find a subcollection $\mathcal{S}' \subseteq \mathcal{S}$, so that the sum of the numbers of CCs of the formed clusters $\sum_{C\in \mathcal{S}'} (f(C)\cdot|C|)$ is maximized.
Besides, the formed clusters should satisfy the requirements on cluster sizes, for example, all the formed clusters should have the same desired size $\delta$, or the sizes should be within a scope $\langle\delta_1, \delta_2\rangle$, where $\delta, \delta_1, \delta_2\in \mathbb{Z}^+$, and $\delta_1 < \delta_2$.
According to the new definition of the number of CCs in a cluster, the singleton clusters in $\mathcal{S'}$ don't contribute to the value to be maximized.

A clustering solution is a subcollection $\mathcal{S}' \subseteq \mathcal{S}$, which satisfies the following conditions, $\bigcup_{C_i\in \mathcal{S}'} C_i = \mathcal{N}$ and $C_{i'}\cap C_i =\emptyset$ when $C_{i'}, C_i\in \mathcal{S}'$ and $i'\neq i$.

% when $|C|>1$, and $f(\cdot)=0$ when $|C|=1$.
%As to $f(\cdot):z\rightarrow z$, under this problem setting, if only number of common channels are given consideration, only singleton clusters will be preferred, which contradicts to our goal of clustering CR nodes together, thus we choose $f(\cdot)= |K_{C}|\cdot |C|$, the product of number of common channels and cluster size.
\end{mydef}

%The decision version of \textit{weighted exact cover problem}: 
%Given an universe $U$, and collection $S=\{s_1, s_2, \ldots, s_m\}$ where each subset $s_i\subseteq U$ and is given a weight $w_i$, whether there exists a collection of subsets $\mathcal{C}$ and constant number $\lambda$, so that the union of $\mathcal{C}$ equals to $U$, $s_j\cap s_{j'} = \emptyset$ for different $j$ and $j'\in \{1,2,\ldots, m\}$, and $\sum_{i\in J} w_i \geq \lambda$.

\section{Centralized Solution for Robust Clustering}
\label{centralized_opt}

Based on Definition~\ref{def_centralized_clustering}, the decision version of this problem is to determine whether there exists a non-empty $\mathcal{S}'\subseteq \mathcal{S}$, so that $\sum_{C\in \mathcal{S}'} f(C) \geqslant \lambda$ where $\lambda$ is a real number and $C$ is a cluster.
We have the following theorem on complexity.

\begin{theorem}
\label{theorem1}
Robust clustering problem in CRN is NP-hard, when the maximum size of clusters is larger than 3.
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is in Appendix~\ref{proof_theorem1}.
%This theorem actually indicates that the robust clustering problem in CRN is NP-hard.

As the robust clustering problem in CRN is NP hard, there is no polynomial solution available.
In this section, we will propose a centralized solution which approaches to the global optimal.
The result will be used as a comparison with the proposed distributed schemes. %looks for the solution which approaches to the global optimum, and the latter is the solution which is more suitable for the practical implementation.
%
%Exact cover problem can be solved with Knuth's Algorithm X~\cite{dancingLinks_Knuth} as it finds out all the instances of exact cover, then we can choose the one with the biggest sum of weights. 
%To make the formulation possible, we adopt heuristics in both the objective function and the constraints, besides, the objective is not totally aligned with the object in the Definition~\ref{def_centralized_clustering}.
The problem by the Definition~\ref{def_centralized_clustering} is formulate into an optimization problem~\ref{centralized_opt} which is a binary linear programming problem.

%This example indicates the chose of $\mathcal{C}$ plays an important role on the resultant clustering strategy.
%Meanwhile, it provide a chance to constrain the cluster size by putting groups with desired sizes into $\mathcal{C}$.

%the maximum size of $S$ is the \textit{Bell number} of $N$, and $S$ contains the conditions cluster .
%In this case, the resultant $\mathcal{C}$ is composed with all the singleton clusters, \ie, the cluster which contains one node, and the objective is -38.

%it is possible that there doesn't exist combination of clusters with the same cluster size.
%We thus list all possible clusters whose sizes are from 2 to one certain number \footnote{this number of decided by the density of CR network, along with the occupation of PUs. We set this number as cluster size of the biggest cluster ever appears when conducting distributed schemes.}, and check each combination of clusters to find the best covering of network on the aspect of number of ICCs per cluster.
%The complexity of computation is thus \bigO$(N^\delta)$, $\delta$ is the preferred cluster size.

%The global optimal clustering scheme with respect to the number of common channels is investigated to show the gap with the distributed schemes.

%We apply this centralized scheme on a network with network size $N$ and cluster size $\delta$.
%There is $N\mod \delta=0$, and the expected number of clusters is $C = N/\delta$.
%These tailored parameters don't harm the validity of the performance gap between the two schemes.

\begin{equation}
\begin{aligned}
     &\min\limits_{x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^G (-x_{ij}q_{ij} + (1-w_i)*p) \\
     &\text{subject to}   && \Sigma_{i=1}^G x_{ij} = 1, for \forall j=1, \ldots, N \\
   &&& \Sigma_{j=1}^N x_{ij} = |C_i|*(1-w_i), for \forall i=1, \ldots, G \\
   &&& \text{$x_{ij}$ and $w_i$ are binary variables.}\\
   &&& i\in \{1,2, \cdots G\}, \hspace{0.3cm} j\in \{1,2,\cdots N\}
%\notag
\end{aligned}
\label{centralized_opt}
\end{equation}

The objective is to maximize the sum of the products of the size and the corresponding number of CCs of all the clusters.
%The second component is designed to help generate the clusters whose sizes are the desired cluster size $\delta$.
We minimize the opposite of the real objective to make this problem solvable.
The constraints guarantee to obtain the clusters which together include all the CR users and don't overlap.

The variables of this problem are $x_{ij}$ where $i\in \{1,2,\cdots,G-1, G\}$ and $j\in \{1,2,\cdots,N-1, N\}$.
$N$ is the total number of CR users in network $\mathcal{N}$.
$G = |\mathcal{G}|$ and $\mathcal{G}$ is the collective of clusters in a CRN $\mathcal{N}$, there is $\mathcal{G} \subseteq \mathcal{S}$, where $\mathcal{S}$ is the collection of all the possible clusters as described in Definition~\ref{def_centralized_clustering}.
The subscript $i$ indicates the $i$th cluster in $\mathcal{G}$, and $j$ corresponds to the node ID of a CR node. 
$x_{ij}$ indicates whether the CR node $j$ resides in the $i$th potential cluster, \ie $x_{ij}=1$ means node $j$ resides in the cluster $C_i$. 
%
%Given a CRN $\mathcal{N}$ and desired cluster size $\delta$, we obtain a collection of clusters $\mathcal{G}$ which contains all the \textit{potential} clusters, and the sizes of these clusters are $1,2,\ldots,\delta$.
%Potential clusters are the clusters which satisfy the conditions introduced in Section~\ref{sec:model}. 
%Note that the potential clusters include the singleton ones.
%By permitting the existence of the singleton clusters, we can ensure that the $\mathcal{S}'$ in Definition~\ref{def_centralized_clustering} is always be feasible.
Based on the knowledge on $\mathcal{G}$, we construct a $G\times N$ matrix $Q_{G\text{x}N}$ which is shown in Figure~\ref{costant_matrix_Q}. 
Constant $q_{ij}$ is the element of $Q_{G\text{x}N}$, and its subscripts correspond to the $i$th cluster and CR node $j$ respectively.
$q_{ij}= |K(C_i)|$ when there is $j\in C_i$, and $q_{ij}= 0$ when there is $j\notin C_i$.
In other words, each non-zero element $q_{ij}$ means the number of CCs of the cluster $i$ where node $j$ resides.

\begin{figure}[ht!]
\centering
\bordermatrix{~ 		& 1 	& 2 	& 3 	& \cdots & j & \cdots	& N-1 	& N	\cr
                  1 	& |K(C_1)| 	& |K(C_1)| 	& 0 	& \cdots & \cdots &\cdots	& 0 	& 0	\cr
                  2 	& |K(C_2)| 	& 0 	& |K(C_2)| 	& \cdots & \cdots & \cdots 	& 0 	& 0	\cr
				\vdots  	&\vdots & 	 	& 		&  \vdots		& 		& \vdots \cr
				i 	& 0 	& |K(C_i)| 	& 0 	& \cdots  & \cdots & \cdots 	& |K(C_i)| 	& 0	\cr
				\vdots  	&\vdots & 	 	& 		&  \vdots & \cdots & \vdots 		& 		& \vdots \cr
				\vdots 	& 0  	& 0 	& 0 	& \cdots & \cdots & \cdots 	& |K(C_i')| 	& 0	\cr
				G  	& |K(C_G)| & \cdots	 	& 		&  \vdots	& \cdots & \vdots& 		& \vdots \cr}	
\caption{An example of Matrix $Q$, its rows correspond to all possible clusters, and columns correspond to the CR nodes in the CRN. }
\label{costant_matrix_Q}
\end{figure}

Now let us have a look at the second component in the objective function.
%Although the related work SOC~\cite{LIU_TMC11_2} adopts the product of cluster size and number of CCs as its sole metric, SOC still generates a large number of singleton clusters and a few large clusters.
To discourage the optimization algorithm from generating the clusters which don't have the preferred size, the second item compromises the first component in the objective function when those clusters are generated.
$w_i$ is the auxiliary binary variable which denotes whether the $i$th  cluster $C_i$ is chosen by the optimization algorithm or not.

$$
w_i = \left\{ \begin{array}{rl}
0 &\mbox{if $i$th possible cluster $C_i$ is chosen} \\
1 &\mbox{if $i$th possible cluster $C_i$ is not chosen} \\
\end{array} \right.
$$
When $w_i$ is zero, it means $C_i$ is chosen by the optimization algorithm, in this case, the second item in the objective function becomes $p$, which is defined as follows,

$$
p = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\alpha_1 &\mbox{if $|C_i|=\delta-1$ or $|C_i|=\delta+1$} \\
\alpha_2 &\mbox{if $|C_i|=\delta-2$ or $|C_i|=\delta+2$} \\
\dots
\end{array} \right.
$$
where $\alpha_1, \alpha_2, \cdots$ are positive values, then when $w_i$ is zero ($C_i$ is chosen by the optimization algorithm) and its size is not the desired size $\delta$, the second component will be greater than zero, which will compromise the first component.
In particular, $\alpha$ increases with the divergence between $|C_i|$ and $\delta$ \ie these is $\alpha_2> \alpha_1>0$.
%Because of $w_i$, any cluster which appears in the clustering solution ($w=0$) will compromise the first component.
%Only when that cluster's size is exactly the preferred size $\delta$, the punishment is zero.
%In contrary, when the chosen cluster's size diverges from $\delta$, the objective function suffers \textit{loss}.
%Then it is easy to notice that when the optimization algorithm adopts singleton clusters, the first xx.
%With such design, the solution is encouraged not to generate singleton clusters.
%doesn't follow the definition of $f(C)$ in Definition~\ref{def_centralized_clustering} strictly, where $f(C_i)=0$ when $|C_i|=1$, but our design echoes the definition by exerting the most severe punishment on the singleton clusters in the clustering solution.
%Choice of $\alpha_i$ affects the resultant clusters.

As to the constraints, the first constraint regulates that each CR node should reside in exactly one cluster.
The second constraint regulates that when the $i$th possible cluster $C_i$ is chosen, there will be exactly $|C_i|$ CR nodes residing in the cluster $C_i$.

%\begin{figure}[ht!]
%\centering
%\includegraphics[width=0.45\linewidth]{example.JPG}
%\caption{Example of Matrix Q in a 6-node network, cluster size is set as 2}
%\label{xx}
%\end{figure}

%\begin{figure}[ht!]
%  \centering
%  \includegraphics[width=0.5\linewidth]{figure5final.pdf}
%  \caption{Final cluster formation.}
%  \label{fig4}
%\end{figure}

%Note that the proposed centralized solution is heuristic.
%We reiterate the reasons for pursuing the heuristic scheme, 
%Note the collection of potential clusters is dependant on the network topology and spectrum availability in the network, thus to each specific CRN, the space of solution is different.

\section{Distributed Clustering Algorithm: ROSS}
\label{ross}
The centralized robust clustering scheme is inherently not suitable for the CRN which involves dynamics in multiple aspects.
Whenever clustering or re-clustering is necessary, which could be triggered by either the mobility of CR users or the variance of the available spectrum, the centralized entity needs to collect the information from all the CR nodes, and then distributes the clustering solution after computation.

In this section we will introduce the distributed clustering scheme ROSS.
With ROSS, CR nodes form clusters based on the proximity of the available spectrum in their neighborhood. 
ROSS consists of two cascaded phases: \textit{cluster formation} and \textit{membership clarification}.

%is based on the spectrum sensing results in its neighbours , then

\subsection{Phase I - Cluster Formation}
\label{phaseI}
After conducting spectrum sensing and communicating with neighbors, every CR node is aware of the channels which are available on themselves and also those on their one hop neighbors.
Two metrics are proposed to characterize the proximity in terms of available spectrum between CR node $i$ and its neighborhood:

\begin{itemize}

\item \textit{Individual connectivity degree} $d_i$: $d_i=\sum_{j\in \text{Nb}(i)}\vert K_i\cap K_j\vert$. 
$d_i$ is the total number of the CCs between node $i$ and every its neighbor.
It is an indicator of node $i$'s adhesion to the CRN.

\item \textit{neighborhood connectivity degree} $g_i$: $g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. 
It is the number of CCs which are available for $i$ and all its neighbors.
$g_i$ represents the ability of $i$ to form a robust cluster with its neighbors.
\end{itemize}
Individual connectivity degree $d_i$ and neighborhood connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where every node's connectivity vector is shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph of the example CRN and the connectivity vector $(d_i, g_i)$ for each node. The sets of the indices of the available channels sensed by each node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. Dashed edge indicates the end nodes are within each other's transmission range.}
	\label{fig1}
\end{figure}

After introducing the connectivity vector, we proceed to introduce the first phase of algorithm ROSS.
In this phase, cluster heads are determined through a series of autonomous negotiations among the CR users, then the neighborhoods of the cluster heads become clusters.

\subsubsection{Determining Cluster Heads and Form Clusters}
In this phase, each CR node decides whether it is a cluster head by comparing its connectivity vector with its neighbors.
When CR node $i$ has lower individual connectivity degree than all its neighbors except for those which have already identified to be cluster heads, node $i$ becomes a cluster head.
%In other words, CR node $i$ becomes cluster head if $d_i > d_k, \forall k\in \texttt{Nb}_i\setminus CHs$ ($CHs$ donate the cluster heads existing in $\texttt{Nb}_i$).
If there is another CR node $j$ in its neighborhood which has the same individual connectivity degree as $i$, \ie $d_j = d_i$ and $d_j < d_{k}, \forall k\in \text{Nb}(j)\setminus \{\text{CHs}\cup i\}$, then the node between $i$ and $j$, which has higher neighborhood connectivity degree will become the cluster head, and the other node become one member of the newly identified cluster head. 
If $g_i = g_j$ as well, the node ID is used to break the tie, \ie the one with smaller node ID becomes a cluster head.
%
The node which is identified as a cluster head broadcasts a message to notify its neighbors of this change, and its neighbors which are not cluster heads become cluster members
\footnote{The reasons for the occurrence of the cluster heads in the neighborhood of a new cluster head will be explained in Section~\ref{ross_p1_guarantee_ccc} and \ref{ross_p2_cluster_pruning})}.
The pseudo code for the cluster head decision and the initial cluster formation is shown in Algorithm~\ref{alg0} in the appendix.

After receiving the notification from a cluster head, a CR node $i$ is aware that it becomes a member of a cluster. 
Consequently, $i$ sets its individual connectivity degree to a positive number $M > |\mathcal{K}| \cdot N$, and broadcasts the new individual connectivity degree to all its neighbors. 
%wrong:
%Note that there is no cluster head which receives the notification on cluster head eligibility, and there is a Lemma on this.
%\begin{theorem}{lemma}
%There is no cluster head receiving the notification message on cluster head eligibility, which is sent from a node which newly becomes a cluster head.
%\end{theorem}
%\begin{proof}
%We use contradiction to prove. 
%Assume a CR node $j$ is cluster head, and receives a notification on cluster head eligibility form CR node $i$. 
%According to the assumption on the reciprocal communication link, $i$ should have received the notification message from $j$ when $j$ becomes cluster head.
%Then CR node $i$ will set its individual connectivity degree to zero, and has no possibility to become a cluster head.
%\end{proof}
When a CR node $i$ is associated to multiple clusters, \ie $i$ has received multiple notifications of cluster head eligibility from different CR nodes, $d_i$ is still set to be $M$. 
%In other words, with this manipulation on its individual connectivity degree, every CR node in the network will be involved in the clustering process, the proof will be given later.
The manipulation of the individual connectivity degree of the cluster members fastens the speed of completing choosing the cluster heads.
%, which enables the nodes which are outside of the cluster can possibly become cluster heads or can also be included into other clusters.
We have the following theorem to show that as long as a secondary user's individual connectivity degree is greater than zero, every secondary user will eventually be either integrated into a certain cluster, or becomes a cluster head.
\begin{theorem}
\label{clustering:theorem}
Given a CRN, it takes at most $N$ steps that every secondary user either becomes cluster head, or gets included into at least one cluster.
\end{theorem}
Here, by \textit{step} we mean one secondary user executing Algorithm~\ref{clustering:theorem} for one time.
The Proof is in Appendix~\ref{proof_clustering:theorem}.

The procedure of the proof also illustrates the time needed to conduct Algorithm~\ref{clustering:theorem}. 
Consider an extreme scenario, where all the secondary nodes sequentially execute Algorithm~\ref{alg0}, \ie they constitute a list as discussed in the example in the proof.
%
If one step can be finished within certain time $T$, then the total time needed for the network to conduct Algorithm~\ref{clustering:theorem} is $N*T$.
In other scenarios, as Algorithm~\ref{alg0} can be executed concurrently by secondary users which locate in different places, the needed time can be considerably reduced.
%According to Theorem~\ref{clustering:theorem}, Phase I completes within a reasonable amount of time.
%
%This judgement is conducted periodically, and phase I ends after every node ascertains it is cluster head or not.
Let us apply Algorithm~\ref{alg0} to the example shown in Figure~\ref{fig1}.
Node $B$ and $H$ have the same individual connectivity degree, i.e., $d_B=d_H$. As $g_H=2>g_B=1$, node $H$ becomes the cluster head and cluster $C_H$ is $\{H, B, A, G\}$.

\subsubsection{Guarantee the Existence of Common Channels in Every Cluster}
\label{ross_p1_guarantee_ccc}
%After deciding itself being cluster head, CR node broadcasts to notify its neighbours on the dedicated control channel, meanwhile, $i$'s initial cluster is formed immediately, which is $i$'s neighbourhood except for those nodes which have become cluster heads, \ie $C_i=(\text{Nb}_i\setminus \text{CHs})\cup i$.
After executing Algorithm~\ref{alg0}, certain formed clusters may don't possess any CCs.
As decreasing cluster size increases the CCs within a cluster, for those clusters having no CCs, certain nodes need to be eliminated to obtain at least one CC.
The sequence of elimination is performed according to an ascending list of nodes which are sorted by the number of common channels between the nodes and the cluster head. 
In other words, the cluster member which has the least common channels with the cluster head is excluded first.
If there are multiple nodes having the same number of common channels with the cluster head, the node whose elimination brings in more common channels will be excluded.
If this criterion meets a tie, the tie will be broken by deleting the node with smaller node ID.
It is possible that the cluster head excludes all its neighbors and resulting in a singleton cluster which is composed by itself.
%At the end of this procedure every formed cluster has at least one common channel.
The pseudo code for cluster head to obtain at least one common channel is shown in Algorithm~\ref{alg_size_control_available_CCC}.
As to the nodes which are eliminated from the previous clusters, they restore their original individual connectivity degrees, executes Algorithm~\ref{alg0}, and become either cluster heads or get included into other clusters afterwards according to Theorem~\ref{clustering:theorem}.

During Phase I, when ever a CR node is decided to be a cluster head and accordingly forms a cluster, or its cluster's composition is changed, the cluster head will broadcast the updated information about its cluster, which includes the sets of available channels on all its cluster members.

\subsubsection{Cluster Size Control in Dense CRN}
\label{ross_p2_cluster_pruning}

In this subsection, we illustrate the pressing necessity to control the cluster size when CRN becomes denser.

We consider a cluster $C(i)$ where $i$ is CH in a dense CRN. 
To make the analysis easier, we assume there is no CHs which are generated within $i$'s neighborhood during the procedure of guaranteeing CCs.
Assuming the CR users and PUs are evenly distributed and PUs occupy the licensed channels randomly, then both CR nodes density and channel availability in the CRN can be seen to be spatially homogeneous.
In this case the formed clusters are decided by the transmission range and network density.
According to Algorithm~\ref{alg0}, the nearest cluster heads could locate just outside node $i$'s transmission range.
%
An instance of this situation is shown in Figure~\ref{clusters_denseNetwork}.
In the figure, black dots represent cluster heads, the circles denotes the transmission ranges of cluster heads.
Cluster members are not shown in the figure.
%Circles represent the transmission range of cluster head, within which CR nodes are absorbed in cluster.
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.3\linewidth]{clusters_denseNetwork_2.pdf}
  \caption{Clusters formation in extremely dense CRN. Black dots are cluster heads, cluster members are not drawn.}
  \label{clusters_denseNetwork}
\end{figure}
Let $l$ be the length of side of simulation plan square, and $r$ be CR's transmission radius.
Based on the aforementioned analysis and geometry illustration as shown in Figure~\ref{clusters_denseNetwork}, we give an estimate on the maximum number of generated clusters, which is the product of the number of cluster heads in one row and that number in one line, $l/r * l/r = l^2/r^2$.
%Now we verify the estimation with simulation.
%We distribute CR and primary users randomly on a square plain, $r$=10 and $l$=50.
%Network density is increased by adding more CR users.
%We now have a look at how does the network density affect the cluster size when the transmission range is constant.
%This implicates when the cluster size is decided by the density of the network.
%As to SOC, the membership of one cluster is decided after a complex process, and the cluster size is roughly the same with one neighborhood.xxxx
%We can see from the example that although two neighbouring clusters can overlap greatly with each other, no cluster head will be covered by other clusters.
Given $r$=10 and $l$=50, the maximum number of clusters is 25.
The number of clusters in the simulation is shown in Figure~\ref{number_clusters_scale}.
Simulation is run for 50 times and the confidence interval is 95\%.
With the increase of CR users, network density (the average number of neighbours) increases linearly, and the number of clusters approaches to 25 which complies with the estimation.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\linewidth]{number_clusters_upperBound.pdf}
  \caption{The correlation between the number of formed clusters and network density.}
  \label{number_clusters_scale}
\end{figure}

Both analysis and simulation show that when applying ROSS, after the number of clusters saturates with the increase of network density, the cluster size increases linearly with the network density, thus certain measures are needed to curb this problem.
This task falls to the cluster heads.
To control the cluster size, cluster heads prune their cluster members to achieve the desired cluster size.
The desired size $\delta$ is decided based on the capability of the CR users and the tasks to be conveyed.
Given the desired size $\delta$, a cluster head excludes members sequentially according to the following principle, the absence of one cluster member leads to the maximum increase of common channels within the cluster.
This process ends when the size of resultant cluster is $\delta$ and at least one CC is available.
This procedure is similar with that to guarantee the existence of CCs in cluster, thus the algorithm can reuse Algorithm~\ref{alg_size_control_available_CCC}.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.5\linewidth]{figure2.pdf}
  \caption{Clusters formation after the phase I of ROSS. CR nodes $A, B, D$ are debatable nodes as they belong to more than one cluster.}
  \label{fig2}
\end{figure}

\subsection{Phase II - Membership Clarification}
\label{membershipClarification}
%\subsubsection*{Problem Description}
After applying phase I of ROSS to the example in Figure~\ref{fig1}, the resulted clusters are shown in Figure~\ref{fig2}.
We notice that nodes $A, B, D$ are included in more than one cluster. 
We refer these nodes as \textit{debatable nodes} as their cluster affiliations are not decided.
The clusters which include the debatable node $i$ are called \textit{claiming clusters} of node $i$, and the set of these clusters is denoted as $S_i$.  
%Actually, debatable nodes extensively exist in CRN with larger scale.
The debatable nodes which are generated from the first phase of ROSS should be exclusively associated with only one cluster and be removed from the other claiming clusters, this procedure is called \textit{cluster membership clarification}.
%We will introduce the solution for cluster membership clarification in the following.

%In particular, the un-affiliation of one debatable nodes from a cluster increasing the set $K_C$ of ICCs of cluster $C$ (at the cost of potentially decreasing $R_C$, the set of OCCs).

%\newtheorem{observation}{Observation}
%\label{observation}
%\begin{observation}
%If the number of nodes within a cluster decreases, the number of common channels will increase or keep constant.
%\end{observation}
%
%\begin{proof}
%Contradiction, To be continued
%\end{proof}
%From observation 1 we know that the procedure of membership clarification will increase the set of common channels for some clusters and accordingly strengthen the robustness of intra connectivity.

% % % % %	dependancy!
%An debatable node belongs to multiple clusters, and in the same time, it is possible that several debatable CR nodes locate within one same cluster. %Each debatable node tries to increase the sum of ICC of the clusters which it belong to. More specifically, 
%For a debatable node $i\in S_i$ after phase I, as to clarify its membership, it will choose one cluster $C\in S_i$ to stay and withdraw from the other clusters in $S_i$ with the consideration of increasing ICCs within $S_i$ by the largest margin.

\subsubsection{Distributed Greedy Algorithm (DGA)}
%Debatable node $i$ is aware of all its claiming clusters in $S_i$. 
Assume a debatable node $i$ needs to decide one cluster $C\in S_i$ to stay, and thereafter leaves the rest others in $S_i$.
In this process, the principle for $i$ is that its move should result in the greatest increase of CCs in all its claiming clusters.
%The set of available channels of one cluster are known by the debatable nodes which locate in that cluster. %this is finished 
Note that node $i$ is aware of the spectrum availability on all the cluster members of each claiming cluster, thus node $i$ is able to calculate how many more CCs can be produced in one claiming cluster if $i$ leaves that cluster.
%Based on this calculation, $i$ decides on one claiming cluster to stay and leaves the other claiming clusters.
If there exists one cluster $C\in S_i$, when $i$ leaves this cluster brings the least increased CCs than leaving any other claiming clusters, then $i$ chooses to stay in cluster $C$.
When there comes a tie, among the claiming clusters, $i$ chooses to stay in the cluster whose cluster head shares the most CCs with $i$.
In case there are multiple claiming clusters demonstrating the same on the aforementioned metric, node $i$ chooses to stay in the claiming cluster which has the smallest size.
Node IDs of cluster heads will be used to break tie if all the previous metrics could not decide on the unique claiming cluster for $i$ to stay.
The pseudo code of this algorithm is given as Algorithm~\ref{alg4}.
%To conduct Algorithm~\ref{alg4}, debatable node $i$ needs to know the necessary information about its claiming clusters, \ie $K(C)$ (the set of available channels in cluster $C$), $K(h_C)$ (the set of available channels on $C$'s cluster head $h_C$) and $|C|, C\in S_i$ (sizes of $i$'s claiming clusters).
After deciding its membership, debatable node $i$ notifies all its claiming clusters of its choice, and the claiming clusters from which node $i$ leaves also broadcast their new cluster composition and the spectrum availability on all their cluster members.

The autonomous decisions made by the debatable CR nodes raise the concern on the endless chain effect in the membership clarification phase.
A debatable node's choice is dependent on the compositions of its claiming clusters, which can be changed by other debatable nodes' decisions.
As a result, the debatable node which makes decision first may change its original choice, and this process may go on forever.
To erase this concern, we formulate the process of membership clarification into a game, where a equilibrium is reached after a finite number of best response updates made by the debatable nodes.

%%  update their choices based on other debatable nodes' choices, and this process 
%%%thesis needs1
%For example, assuming one debatable node $i$ locates in cluster $C\in S_i$, and $C$ has more than one debatable node except for $i$.
%Assuming node $i$ decides to stay in cluster $C$, then one another debatable nodes $j$ decides its affiliation, and there is $j\in C\in S_i$.
%$j$'s move decreases cluster $C$'s size, and could possibly trigger node $i$ to alter its previous decision to leave $C$, as $C$'s size is smaller now and leaving it may result in more increase of CCs in $S_i$.
%At this point of time, debatable node $j$ may rejoin cluster $C$ due to the changes in $S_j$, then both node $i$ and $j$ face the \textit{same} situation again.
%%\footnote{Actually it is not totally same as before, as there are some new changes within $S_j$.}
%%To maximize the increase of common channel in clusters in $S_j$, $j$ either stays in $C$, or leaves cluster $C$ and stay another cluster in $S_j$.
%%$j$'s decision changes $C$'s membership, which will affect $i$'s decision.
%%Assume $i$ makes decision before $j$ and chooses to stay in $C$, as which brings the most common channels in $i$' claiming clusters, and then it is $j$' turn to choose cluster.
%%If $j$ leaves $C$, the smaller cluster $C$ will possibly make $i$ to leave it and join another cluster.
%%and the choice of $j$ to stay in $C$ or not possibly changes $C$'s membership and  which potentially further triggers node $i$ to alter its previous decision. 
%Thence, we need to answer this concern before implementing ROSS-DGA.
%%, and if it converges, how good such a distributed scheme performs.

\subsubsection{Bridging ROSS-DGA with Congestion Game}
\label{clustering:phaseII:game}
Game theory is a powerful mathematical tool for studying, modelling and analysing the interactions among individuals.
A game consists of three elements: a set of players, a selfish utility for each player, and a set of feasible strategy space for each player. In a game, the players are rational and intelligent decision makers, which are related with one explicit formalized incentive expression (the utility or cost).
Game theory provides standard procedures to study its equilibriums~\cite{game_for_communication_01}.
In the past few years, game theory has been extensively applied to problems in communication and networking~\cite{Neel06analysisand, Wang_gtc_crn_survey_2010}.
Congestion game is an attractive game model which describes the problem where participants compete for limited resources in a non-cooperative manner, it has good property that Nash equilibrium can be achieved after finite steps of best response dynamic, \ie each player choose strategy to maximizes/minimizes its utility/cost with respect to the other players' strategies.
Congestion game has been used to model certain problems in internet-centric applications or cloud computing, where self-interested clients compete for the centralized resources and meanwhile interact with each other.
For example, server selection is involved in distributed computing platforms~\cite{Cloud_Computing_2010}, or users downloading files from cloud, etc.

To formulate the debatable nodes' membership clarification into the desired congestion game, we observe this process from a different (or opposite) perspective. 
From the new perspective, the debatable nodes are regarded to be isolated and don't belong to any cluster, in other words, their claiming clusters become clusters which are beside them. 
Now for the debatable nodes, the previous problem of deciding which clusters to leave becomes a new problem that which cluster to join.
In the new problem, debatable node $i$ chooses one cluster $C$ out of $S_i$ to join if the decrease of CCs in cluster $C$ is the smallest in $S_i$, and the decrease of CCs in cluster $C$ is $\sum_{C\in S_i}\Delta\vert K(C) \vert=\sum_{C\in S_i}({\vert K(C) \vert-\vert K(C\cup i) \vert})$.
The interaction between the debatable nodes and the claiming clusters is shown in Figure~\ref{debatable_nodes_claiming_cluster}.
%The concern on convergence appears again as we have discussed in the previous subsubsection.
%We give proof on convergence under game theoretic framework.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.25\linewidth]{singletongame_png.png}
  \caption{Debatable nodes and claiming clusters}
  \label{debatable_nodes_claiming_cluster}
\end{figure}

%In the following we will introduce an \textit{server matching}~\cite{kothari:congestion_serverMatching} problem to illustrate congestion game's application in communication systems.

In the following, we show that the decision of debatable nodes to clarify their membership can be mapped to the behaviour of the players in a \textit{player-specific singleton congestion game} when proper cost function is given.
The game to be constructed is represented with a 4-tuple $\Gamma=(\mathcal{P},\mathcal{R},\sum_{i, i \in \mathcal{P}}, f)$, and the elements in $\Gamma$ are explained below,
%To make the model of this game more clear, we make some change to our original problem. Previously, the nodes in overlapping areas belong to more than one cluster, and our scheme is to remove them out of some clusters to increase the set of common channels within the cluster form which the mode leave. In the new model, we assume all the nodes in overlapping nodes don't belong any cluster and the problem become into how do these nodes decide which cluster to join.

%The components of the game are listed as follows,

\begin{itemize}
%	\item $\mathcal{D}=\left\{1,\ldots,n\right\}$, the set of players (debatable nodes).
	\item $\mathcal{P}$, the set of players in the game, which are the debatable nodes in our problem.
%	\item $\mathcal{R}=\left\{1,\ldots,m\right\}$, the set of resources which player can choose, which are all the clusters in our model.
	\item $\mathcal{R} = \cup S_i, i\in \mathcal{P}$, denotes the set of resources for players to choose, in our problem, $S_i$ is the set of claiming clusters of node $i$, and $\mathcal{R}$ is the set of all claiming clusters.
	\item Strategy space $\sum_i, i \in \mathcal{P}$ is the set of claiming clusters $S_i$.
	As debatable node $i$ is supposed to choose only one claiming cluster, then only one piece of resource will be allocated to $i$.%, accordingly this congestion game is a singleton game.
	%when $i$ makes decision, only one resource (one claiming cluster) from the allowed resources is allocated.

	%\item We denote by $\mathcal{S}=\left(\mathcal{S}_1,\ldots,\mathcal{S}_n\right)\in \sum_1\times \cdots\times\sum_n$ the state of game where player $i$ plays strategy $\mathcal{S}_i\in \Sigma_i$.

	\item %For the clusters which are possible destination of debatable nodes, the decrease of common channels caused by different debatable node' join can be different because of the heterogeneity of channel availability within itself and on the debatable nodes. %furthermore, the sequence of debatable node's join can also alter the decrease. 
	The utility (cost) function $f(C)$ as to a resource $C$. 
	$f(C) = \Delta\vert K^i(C)|, C\in S_i$, which represents the decreased number of CCs in cluster $C$ when debatable node $i$ joins $C$.
	As to cluster $C\in S_i$, the decrease of CCs caused by including the debatable nodes is $\sum_{i:C\in S_i, i\rightarrow C} \Delta\vert K^i(C) \vert$. 
$i\rightarrow C$ means $i$ joins cluster $C$.
Obviously this function is non-decreasing with respect to the number of nodes joining cluster $C$.

The utility function $f$ is not purely decided by the number of players accessing the resource (debatable nodes join claiming clusters), which happens in a canonical congestion game.
The reason is in this game the channel availability on debatable nodes is different.
Given two same groups of debatable nodes and their sizes are the same, when the nodes are not completely the same (neither are the channel availabilities on these nodes), the cost happened on one claiming cluster could be different if the two groups of debatable nodes join that cluster respectively.
%In a canonical congestion game, the cost (or pay off) is function of only the number of players occupying the resource, and is mono-
%In this new game, the cost function 
Hence, this congestion game is player specific~\cite{Ackermann06purenash}.
In this game, every player greedily updates its strategy (choosing one claiming cluster to join) if joining a different claiming cluster minimizes the decrease of CCs $\sum_{i:C\in S_i} \Delta\vert K^i(C) \vert$, and a player's strategy in the game is exactly the same with the behaviour of a debatable node in the membership clarification phas.

%	\item The Rosenthal's potential function \cite{Rosenthal} of this congestion game is given by:
%	\begin{equation*}
%	\phi(S)=\sum_{C\in\mathcal{R}} \sum_{i:C\in S_i} \Delta\vert K^i_C \vert   	
%   	%\sum_{i=1}^N \Delta^{i}_{p}(S)=\sum_{i=1}^N \sum_{r\in S_i}\Delta^{i}_{r}(t)	
%  	% \Delta =\sum_{i=1}^N w_i (x_i - \bar{x})^2 .
%	\end{equation*}
%All the players in this game greedily update their strategy to minimize the potential function (congestion), this process is exactly the same with the network behaviour under \textit{Distributed Greedy Algorithm}.

%	\item It is an asymmetric game because the sets of strategies shared by different players are different.
%	\item The total cost is: 
%\begin{equation*}
%   \sum_{i=1}^N \Delta^{i}_{p}(S)=\sum_{i=1}^N \sum_{p\in s_i} \Delta^{i}_{p}(n_p(S))
%  % \Delta =\sum_{i=1}^N w_i (x_i - \bar{x})^2 .
%\end{equation*}

%This is the global objective we want to minimize.
\end{itemize}

%Singleton congestion game is a special type of matroid game~\cite{Milchtaich1996111,}. 
%It is known that player-specific matroid congestion game admit pure equilibrium,

As to singleton congestion game, there exists a pure equilibria which can be reached with the best response update, and the upper bound for the number of steps before convergence is $n^2*m$~\cite{Ackermann06purenash}, where $n$ is the number of players, and $m$ is the number of resources.
In our problem, the players are the debatable nodes, and the resources are the claiming clusters.
Thus the upper bound of the number of steps can be expressed as $\mathcal{O}(N^3)$.

In fact, the number of steps which are actually involved in this process is much smaller than $N^3$, as both $n$ and $m$ are considerably smaller than $N$.
The percentage of debatable nodes in $\mathcal{N}$ is illustrated in Figure~\ref{percentage_overlapping_node}, which is between 10\% to 60\% of the total number of CR nodes in the network.
The number of clusters heads, as discussed in Section~\ref{phaseI}, is dependent on the network density and the CR node's transmission range.
As shown in Figure~\ref{number_clusters_scale}, the cluster heads take up only 3.4\% to 20\% of the total number of CR nodes.

%and the number of steps towards \textit{Nash Equilibrium} is upper-bounded\footnote{Here we present this with modifying the original conclusion in \cite{Ackermann06purenash} according to our model.} by $ n^2\cdot m $. In our context, $n$ is the number of debatable nodes, $m$ is number of clusters in CRN, %$rk(\Gamma)$ of the matroid  is the cardinality of the maximal independent sets, which is 1 in the case of singleton game, 
%so the total time complexity to achieve the \textit{Nash Equilibrium} using greedy approach is 	.
%%(XXX Just mention after this complexity result the relationship to the system mdeol XX)
%This is upper-bounded (in the worst case) by $O(\vert I\vert^3)$. 
%Based on above model and analysis, phase II converges is Algorithm~\ref{alg4} is run by debatable nodes. 
%Although the game version of DGA can achieve \textit{Nash Equilibrium}, the whole scheme can possibly obtain sub-optimal result.    %, furthermore,this stable state is a local minimum of the global decrease function.
%\todo[inline]{The number of steps, or the upper bound of steps in convergence needs a formal proof}

\subsubsection{Distributed Fast Algorithm (DFA)}
%The convergence speed of DGA is large recalling that the number of steps is of $\mathcal{O}(N^3)$.
On the basis of ROSS-DGA, we propose a faster version ROSS-DFA which differs from ROSS-DGA in the second phase.
With ROSS-DFA, debatable nodes decide their respective cluster heads once.
The debatable nodes consider their claiming clusters to include all their debatable nodes, thus the membership of claiming clusters is static and all the debatable nodes can make decision simultaneously without considering the change of membership of their claiming clusters.
As ROSS-DFA is quicker than ROSS-DGA, the former is especially suitable for the CRN where the channel availability changes dynamically and re-clustering is necessary.
To run ROSS-DFA, debatable node executes only one loop in Algorithm~\ref{alg4}.

Now we apply both ROSS-DGA and ROSS-DFA to the toy network in Figure~\ref{fig2} which has been applied the phase I of ROSS.
In the network, node $A$'s claiming clusters are cluster $C(C), C(H)\in S_A$, their members are $\{A,B,C,D\}$ and $\{A,B,H,G\}$ respectively. 
The two possible strategies of node $A$ is illustrated in Figure \ref{fig3}.
In Figure \ref{AinC}, node $A$ staying in $C(C)$ and leaving $C(H)$ brings 2 more CCs to $S_A$, which is more than that brought by another strategy showed in \ref{AinH}.
After the decisions made similarly by the other debatable nodes $B$ and $D$, the final clusters are formed as shown in Figure~\ref{final_clustering_ross}.

%Using DFA in phase II, the time complexity is decreased drastically to 1. Thus, the total complexity of ROSS-DFA is $|I|$, while, ROSS-DGA's complexity is $|I|^3$ in the worst case.

\begin{figure}[h]
\centering
\subfigure[Node A stays in cluster $C(C)$, quits $C(H)$, $\Delta\vert K(C(C))\vert+\Delta\vert K(C(H))\vert=2$]{
\includegraphics[width=0.435\linewidth]{figure4AinC.pdf}
\label{AinC}
}
\hspace{.15 in}
\subfigure[Node A stays in cluster $C(H)$, quits $C(C)$, $\Delta\vert K(C(C))\vert+\Delta\vert K(C(H))\vert=1$]{
\includegraphics[width=0.435\linewidth]{figure4AinH.pdf}
\label{AinH}
}
\caption[]{Membership clarification: possible cluster formations caused by node A's different choices} %\subref{node A in $C_C$}, \subref{node A in $C_H$}}
\label{fig3}
\end{figure}

\begin{comment}
As an example, when node A comes to decide which cluster to stay, the memberships of relevant clusters, like $C_C$ and $C_H$, are $\{C,B,D,A\}$ and $\{H,B,G,A\}$ respectively. Before the other two debatable nodes B and D making their belonging clear, cluster $C_C$ and $C_H$ have them in the same time. So node A can decide which cluster to belong to without considering other debatable nodes' action. There are two strategies for node A, which is illustrated in Figure \ref{fig3}. Because staying in cluster $C_C$ brings in more common channels within relevant clusters, node A finally choose cluster $C_C$ to stay and caveat from cluster $C_H$. The membership of $C_H$ is updated in the same time. Node B and D undertake the same process and the clusters are formed finally as Figure \ref{fig4} shows.

Because debatable nodes can conduct membership clarification abased on static membership information of relevant clusters, thus no iteration happens in this process. The time complexity of this algorithm is only decided by the number of debatable nodes, which is maximal $O(\vert I\vert)$.

As an example, when node A comes to decide which cluster to stay, the memberships of relevant clusters, like $C_C$ and $C_H$, are $\{C,B,D,A\}$ and $\{H,B,G,A\}$ respectively. Before the other two debatable nodes B and D making their belonging clear, cluster $C_C$ and $C_H$ have them in the same time. So node A can decide which cluster to belong to without considering other debatable nodes' action. Figure \ref{AinC}. Because staying in cluster $C_C$ brings in more common channels within relevant clusters, node A finally choose cluster $C_C$ to stay and retreat from cluster $C_H$. The membership of $C_H$ is updated in the same time. Node B and D undertake the same process and the clusters are formed finally as Figure \ref{final_clustering_ross} shows.
\end{comment}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\linewidth]{final_clustering_ross.pdf}
  \caption{Final formation of clusters. $K(C(C)),K(C(E)),K(C(H))$ are shown beside corresponding clusters.}
  \label{final_clustering_ross}
\end{figure}

\subsection{Apply the Centralized and Comparison Distributed Schemes in the Example CRN}
After introducing the resulted clusters from ROSS in Fig.~\ref{final_clustering_ross}, we apply the other robust clustering schemes in the same example CRN in Figure~\ref{fig1}.
%We look into how does the centralized scheme perform in the toy example of the CRN in Figure~\ref{fig1}.
%
As to the centralized robust clustering scheme, we let the desired cluster size $\delta$ be 3.
A collection of clusters $\mathcal{G}$ is obtained, which contains all the clusters satisfying the conditions of clusters in Section~\ref{sec:model} and the sizes of clusters are 1, 2 or 3. 
$\mathcal{G}=\{\{A\}, \{B\},\dots,\{B,C\},\{B,A\},\{B,H\},\cdots,\{B,A,C\},$\\$\{B,H,C\}, \{A,D,C\},\cdots\}$, and $G = |\mathcal{G}|=38$.
%
When $\alpha_1$ and $\alpha_1$ are set as 0.2 and 0.8, the formed clusters are shown in Fig.~\ref{fig:final_clustering_LP}.
%$\{\{D,E,F\},\{A,C,G\},\{H,G\}\}$, the numbers of CCs are 2, 3, 3.
The resulted clustering solutions from SOC is shown in Fig.~\ref{fig:final_clustering_soc}.%, clusters are $\{A,B,C,D,G\},\{E,F\},\{H\}$, and the numbers of CCs are 2, 3, 4.
%The solution resulted from ROSS-DGA is $\{\{B,H,G\},\{C,A\},\{D,E,F\}\}$ (in Figure~\ref{fig4}), the numbers of CCs are 2, 4, 2.
%ROSS-DFA generates the same result with ROSS-DGA in this example.

As to the average number of CCs, the results of ROSS (including both ROSS-DGA and ROSS-DFA), centralized and SOC are 2.66, 2.66, and 3 respectively. 
Note there is one singleton cluster $C(H)$ generated by SOC, which is not preferred.
When we take no account of the singleton clusters, then the average number of common channels of SOC drops to 2.5. 
\begin{figure}[ht]
\begin{center}
\subfigure[Resulted from SOC]{\includegraphics[width=0.435\linewidth]{final_clustering_soc}\label{fig:final_clustering_soc}}
\hspace{0.15 in}
\subfigure[Resulted from the centralized clustering scheme]{\includegraphics[width=0.435\linewidth]{final_clustering_LP}\label{fig:final_clustering_LP}}
\end{center}
\caption{Final clusters formed in the example network when being applied with SOC and the centralized clustering scheme.}
\label{fig:final_clustering}
\end{figure}

\section{Performance Evaluation}
\label{performance}
%In this section, we evaluate the performances of ROSS.
The schemes involved in the simulation are listed as follows,
\begin{itemize}
\item ROSS without size control, \ie ROSS-DGA and ROSS-DFA.
\item ROSS-$\delta$-DGA and ROSS-$\delta$-DFA, which are the variants of ROSS with the size control feature.
$\delta$ is the preferred cluster size.
\item SOC~\cite{LIU_TMC11_2}, a distributed clustering scheme pursuing cluster robustness.
\item Centralized robust clustering scheme. 
The formulated optimization is an integer linear optimization problem, which is solved by the function $bintprog$ provided in MATLAB.
\end{itemize}
The authors of~\cite{LIU_TMC11_2} compared SOC with other schemes in terms of the average number of CCs of the formed cluster, on which SOC outperforms other schemes by 50\%-100\%. 
SOC's comparison schemes are designed either for ad hoc network without consideration of channel availability~\cite{Basagni99}, or for CRN but just considering connection among CR nodes~\cite{Zhao07}. 
Thus SOC is chosen to be the only distributed scheme as comparison, besides, we also compare ROSS with the centralized scheme.
%
We investigate the schemes with the following metrics.

\begin{itemize}
\item \textbf{Average number of CCs per non-singleton cluster.}
Non-singleton cluster refers the cluster whose cluster size is larger than 1.
%This metric denotes the robustness of the formed non-singleton clusters.
Comparing with the metric adopted by SOC~\cite{LIU_TMC11_2}, which is the average number of CCs per cluster without excluding the singleton clusters, this metric provides an more accurate description of the robustness of the non-singleton cluster.
Having more CCs per non-singleton clusters means these clusters have higher probability to survive when the primary users' operation becomes more intense.
%which is biased as the singleton clusters (isolated CR users) don't get benefits from the cluster structure.
Although this metric doesn't disclose the information about the CR nodes which are not included in any non-singleton clusters, we still examine this metric as the number of CCs is involved in the utility which is adopted by all robust clustering schemes.
%, thus it fails to give an overview of how many CR nodes can benefit from non-singleton clusters.

%the CR nodes with more channels could be formulated as singleton clusters.
%This happens in SOC solution, whose objective is to improve the average number of common channel over \textit{all} clusters, \ie including the singleton clusters, thus many CR nodes with more channels are turned into clusters.
%As we try to look into the robustness of clusters of CRs, we exclude those singleton clusters.

\item \textbf{Cluster sizes.}
%Specific clusters size is pursued in many applications due to energy preservation and the system design ~\cite{clustering_globecom11}.
The distribution of CRs residing in the formed clusters with different sizes is presented.

\item \textbf{Robustness of the clusters against newly added PUs.}
We increase the number of PUs to challenge the clusters, and count the number of \textit{unclustered} CR nodes which are the synonyms of the singleton clusters.
This metric indicates the robustness of clusters from a more practical point of view, \ie as to the clusters formed for a given CRN and spectrum availability, how many CR nodes can still make use of the clusters when the spectrum availability decreases.
%robustness of clusters, as this metric directly shows how many nodes can make use of the cluster structure.
%When we investigate the performance with moderate and vigorous intensity of primary users' activities, 
%All the robust clustering schemes are proposed for a given CRN, when the available spectrum becomes scarce when the primary users' operation becomes more intense, the metric can illustrate the robustness of clusters by showing the unclustered CR nodes.
%This metric is closely related with cluster robustness, \ie the less CR nodes which are not  nodes means more CR nodes are included into non-singleton clusters which survive the primary users' influence.
%When we vary the intensity of primary users' activity, \eg from low to medium level by increasing the number of primary users, 
%The drawback of this metric is, it doesn't indicate the robustness of the non-singleton clusters when the primary users conquer more spectrum.

\item \textbf{Amount of control messages involved.}
We investigate the number of control messages involved in the clustering process.

\end{itemize}

Simulation consists of two parts, in the first part, we investigate the performance of centralized scheme, and the gap between the distributed schemes and the centralized scheme.
This part is conducted in a small network, as there is no polynomial time solution available to solve the centralized problem.
In the second part, we investigate the performance of the proposed distributed schemes in the CRN with different scales and densities.

We give a brief introduction to the settings which are common for both simulation parts.
CRs and PUs are deployed on a two-dimensional Euclidean plane.
%Complying with the system model, the CR node residing within another CR node's transmission range is seen as neighbour of that CR node.
The number of licensed channels is 10, each PU is operating on each channel with probability of 50\%.
CR users are assumed to be able to sense the existence of primary users and identify available channels.
All primary and CR users are assumed to be static during the process of clustering.
The simulation is written in C++, and the performance results are averaged over 50 randomly generated topologies, and the confidence interval corresponds to 95\% confidence level.

\subsection{Centralized Schemes vs. Decentralized Schemes}
There are 10 primary users and 20 CR users are dropped randomly (with uniform distribution) within a square area of size $A^{2}$, where we set the transmission ranges of primary and CR users to $A/3$.
%There are 10 available channels. 
%With this setting, the average number of neighbours of one CR user is 4.8.
%Each primary user randomly occupies one channel, and 
When clustering scheme is executed, around 7 channels are available on each CR node.
The desired cluster size $\delta$ is 3, the parameters used in the \textit{punishment} for choosing the clusters with undesired sizes are set as follows, $\alpha_1 =  0.4$, $\alpha_2 =  0.6$.

\subsubsection{Number of CCs in Non-singleton Clusters}
\label{ccc_20}
%We first have a look at the average number of common channels per non-singleton cluster.
% shows the average number of the CCs of non-singleton clusters,
%as the singleton clusters (in other words unclustered nodes) don't execute any functionalities of clusters, which are has be discussed in Section~\ref{intro}.
From Figure~\ref{ccc_per_nonsingleton}, we can see the centralized schemes outperform the distributed schemes.
As to the distributed schemes, SOC achieves the best than all the variants of ROSS.
The reason is, SOC is liable to group the neighbouring CRs which share the most abundant spectrum together, no matter how many of them are, thus the number of CC of the formed clusters is higher.
We have discussed the flaw of this metric as it doesn't convey the number of unclustered CR nodes, in fact, SOC generates the most unclustered CRs, which can be seen when we discuss the performance on the number of unclustered CR nodes. 
As to the variants of ROSS, we notice that the greedy mechanism increases CCs in non-singleton clusters significantly.
%, this is due to the procedure that debatable nodes greedily look for better affiliation to improve the number of CCs.
%We also notice that the size control feature doesn't affect the number of CCs for both ROSS-DGA and ROSS-DFA.
%Size control mechanism converts the large clusters into small ones, but meanwhile clusters with the desired sizes have to be made when forming smaller clusters is possible.
%, which can also be observed in the large scale CRN in Section~\ref{largeScaleCRN}.

%This is because the desired cluster size happens to be the average size of clusters generated by ROSS-DGA and ROSS-DFA, then the size control functionality doesn't play effect to increase the number of CCs.

\begin{figure*}[th]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{ccc_20.pdf}\par\caption{Number of common channels of non-singleton clusters}\label{ccc_per_nonsingleton}
    \includegraphics[width=\linewidth]{cdf_clusterSize_20.pdf}\par\caption{Cumulative distribution of CRs residing in clusters with different sizes}\label{size_control}    
    \includegraphics[width=\linewidth]{survival_rate_20.pdf}\par\caption{Number of unclustered CRs with decreasing spectrum availability}\label{singleton_clusters}
\end{multicols}
\caption{Comparison between the distributed and centralized clustering schemes ($N$ = 20)}
\label{compare_dis_centralized}
\end{figure*}

\subsubsection{Cluster Size Control}
\label{cluster_size}
%\begin{figure}[h]
%  \centering
%  \includegraphics[width=0.8\linewidth]{cdf_clusterSize_20.pdf}
%  \caption{Cumulative distribution of CRs residing in clusters with different sizes.}
%  \label{size_control}
%\end{figure}

Figure~\ref{size_control} depicts the empirical cumulative distribution of the CRs residing in certain sized clusters, from which we have two conclusions.
The first, SOC generates more unclustered CR nodes than other schemes.
The centralized schemes produce no unclustered CR nodes, ROSS-DGA/DFA generate 3\% unclustered nodes, as comparison, 10\% of nodes are unclustered when applying SOC.
ROSS-DGA and ROSS-DFA with size control feature generate 5\%-8\% unclustered CR nodes, which is due to the cluster pruning procedure (discussed in section~\ref{ross_p1_guarantee_ccc} and section~\ref{ross_p2_cluster_pruning}).
Second, the centralized schemes and cluster size control mechanism of RPSS generate clusters which satisfy the requirement on cluster size strictly.
%When the desired size is 2, each generated cluster has two members, whereas when the desired size is 3, about 15\% CRs are formed into 2 node clusters.
As to ROSS-DFG and ROSS-DFA with size control feature, CR nodes reside averagely in clusters whose sizes are 2, 3 and 4.
%When ROSS-3-DFA is applied, most number of CRs are in 3 node clusters, nevertheless, slightly less nodes are found in 2 node and 4 node clusters, there are also considerable number of singleton clusters.
%ROSS-3-DGA decreases the clusters sizes and results in more 2 node clusters, the second most CRs are found in 3 node clusters.
The sizes of clusters resulted from ROSS-DGA and ROSS-DFA are disperse, but appear to be better than SOC, i.e., the 50\% percentiles for ROSS-DGA, ROSS-DFA and SOC are 4.5, 5, and 5.5, and the 90\% percentiles for the three schemes are 8, 8, and 9.
%Figure~\ref{size_control} shows distributed clustering schemes are not able to control cluster sizes perfectly, but ROSS-DGA and ROSS-DFA eliminate the clusters whose size diverges largely with the desired one, \ie single node clusters and clusters with size of 13 and 14.

\subsubsection{Robustness of the clusters against newly added PUs}
%When the number of PUs in CRN increases, or their operation becomes more intensive, some clusters don't seize any CCs any more, so that the cluster members and the cluster heads become unclustered, or singleton clusters.
%If there is no common channels available any more because of the new added PRs, the cluster is regarded as destroyed and the former cluster member CRs become unclustered CRs or in other words singleton clusters.
%We investigate the number of singleton clusters with primary users whose intensity of activities are varying.
%We will obtain two observations by examining this metric.
%First, we will know how many CR nodes are unclusterred by applying the robust clustering schemes in a given CRN.
We add PUs in the CRN to decrease the available spectrum, and observe the number of unclustered CR nodes.
Clusters are formed with the presence of 10 PUs in the beginning, then extra 20 batches of PUs are added sequentially, where each batch includes 5 PUs. 
%The transmission range and channel occupancy of the new PU is the same with the previous ones, \ie transmission range is $A/3$, and one channel out of 10 is randomly chosen to operate.

Figure~\ref{singleton_clusters} shows with the increase of PUs, certain clusters disappear and the number of unclustered CR nodes increases.
Among the robust clustering schemes, the centralized scheme with desired size of two generates the most robust clusters, meanwhile, SOC results in the most vulnerable clusters.
The centralized scheme with desired size of three doesn't outperform the variants of ROSS, because the centralized scheme peruses clusters with size of 3 at the expenses of sanctifying CCs.
In contrary, the variants of ROSS generate some smaller clusters which are more likely to survive when PUs' activities become intense.
%The comparison on cluster sizes will be given in Section~\ref{cluster_size}.
% when the number of PRs is 10$\sim$30, when number of PUs is 30$\sim$60, same amount of unclustered CRs are generated with variants of ROSS.
%When there are 75 and more new PRs, centralized scheme with cluster size of 3 results in more unclustered CR nodes than variants of ROSS.
%ROSS with size control distinguishes itself as the size control avoid the appearance of the clusters with large size.

%As to the variants of ROSS, the greedy process in ROSS improves the performance.

%Greedy strategy adopted in the second phase of ROSS improves the robustness of clusters, i.e. ROSS-DGA exceeds ROSS-DFA, and ROSS-$\delta$-DGA surpasses ROSS-$\delta$-DFA.
%When the debatable CRs greedily update their affiliations, one of the metrics is the maximum increase of CCs of the demanding clusters. 
%This observation complies with the result shown in Figure~\ref{ccc_per_nonsingleton}.
%Meanwhile, sizes of more clusters become smaller also contributes more robustness.

%\begin{figure}[h!]
%  \centering
%  \includegraphics[width=0.8\linewidth]{ccc_20.pdf}
%  \caption{Number of common channels for non-singleton clusters, the numbers in the names of schemes annotate the desired cluster size.}
%  \label{ccc_per_nonsingleton}
%\end{figure}
%
%
%\begin{figure}[h]
%  \centering
%  \includegraphics[width=0.8\linewidth]{survival_rate_20.pdf}
%  \caption{Number of CRs which are not included in any clusters}
%  \label{singleton_clusters}
%\end{figure}

\subsubsection{Control Signalling Overhead}
%Different from the clustering schemes proposed in ~\cite{LIU_TMC11_2, clustering_globecom11}, 
%There are two phases for any variants of ROSS.
%Clusters are formed in the first phase, in the second phase, cluster membership is decided so that each node only resides in one cluster.
%Control message exchanges between CR nodes are involved in both phases.

In this section we compare the overhead of signalling involved in different clustering schemes.
%In order to highlight the amount of control signalling only for clustering, 
We omit the control messaged involved in neighbourhood discovery, which is the premise for any clustering scheme.
According to~\cite{complexity_aggregation_2011}, the message complexity is defined as the number of messages used by all nodes.
To have the same metric to compare, we count \textit{the number of transmissions of control messages}, without distinguishing they are sent with broadcast or unicast.
This metric is synonymous with \textit{the number of updates} discussed in Section~\ref{ross}.

As to ROSS, the control messages are generated in both phases.
In the first phase, when a CR node decides itself to be the cluster head, it broadcasts a message containing its ID, cluster members and the set of CCs in its cluster.
%As to ROSS with size control feature, there are same amount of cluster heads with ROSS without enabling size control feature, and the cluster head broadcasts the available channels of the pruned cluster.
In the second phase, a debatable node broadcasts its affiliation to inform its claiming clusters, then the $\text{CH}$s of the claiming clusters  broadcast message about the new cluster members if they are changed due to the debatable node's decision.
The total number of the decisions involved in cluster formation has been analysed in Theorem~\ref{clustering:theorem} and Section~\ref{clustering:phaseII:game} respectively.

Comparison scheme SOC involves three rounds of execution. 
In the first two rounds, every CR node maintains its own cluster and seek to integrate neighbouring clusters, or joins one of them.
The final clusters are obtained in the third round. 
In each round, every CR node is involved in comparisons and cluster mergers.
%Comparing with the second phase of ROSS, only debatable nodes to communicate with cluster heads to clarify their membership.
%The signalling overhead for centralized scheme comes from two processes, the the process of collecting information to the centralized controller, and the process that the controller spreads the clustering result to all the CR nodes.

The centralized scheme is conducted at the centralized control device, but it involves two phases of control message transmission.
The first phase is information aggregation, in which every CR node's channel availability and neighborhood is transmitted to the centralized controller.
The second phase is broadcasting, where the clustering solution is disseminated to every CR node.
%
We adopt the algorithm proposed in~\cite{Efficient_broadcasting_gathering_adhoc} to broadcast and gather information as the algorithm is simple and self-stabilizing.
This scheme needs building a backbone structure to support the communication, and we apply ROSS to generate cluster heads which serve as the backbone, besides, the debatable nodes as used as the gateway nodes between the backbone nodes.
As the backbone is built once and support transmission for multiple times, the messages involved in the clustering process are not counted.
As to the process of information gathering, we assume that every cluster member sends the spectrum availability and its ID to its cluster head, which further forwards or the message to the controller, thus the number of transmission is $N$.
As to the process of dissemination, in an extreme situation where all the gateway and the backbone nodes broadcast, the number of transmission is $h + m$, where $h$ is the number of cluster heads and $m$ is number of debatable nodes.
%, $d$ is the average number of demanding clusters for each debatable node.

The number of control messages which are involved in both ROSS and the centralized scheme is related with the number of debatable nodes.
Figure~\ref{percentage_overlapping_node} shows the percentage of debatable nodes when the CRN network becomes denser, from which we can obtain the value of $m$.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.6\linewidth]{percentage_overlapping_node.pdf}
  \caption{The percentage of debatable nodes after phase I of ROSS.}\label{percentage_overlapping_node}
\end{figure}
%
%As we adopt a simplified communication model, node's transmission is not influenced by collision or interference, 
%Our simulation doesn't consider the behaviour in the physical layer, 
%
%
%Assume we use OSPF~\cite{BCJ10} to aggregate and disseminate information, then the best and worst complexity for is $\mathcal{O}(E)$, where $E$ is the number of edges in the graph which corresponds to the network.
%The minimum number of edges is $n-1$ when the nodes form a line and each node has at more two neighbours, and the maximum number is $N*(N-1)/2$ when the nodes form a complete graph.
%Thus the best message complexity of the centralized scheme is $\mathcal{O}{(N)}$ and the worst is $\mathcal{O}{(N^2)}$.
%
%1. update membership to form X1, 
%2. broadcast new X1, form new X2
%3. broadcast X3
%The complexity parameters are the number of nodes $n$ in network, number of clusters $h$.
The message complexity, quantitative analysis of the number of control messages involved in clustering, and the size of control messages are shown in Table \ref{tab_overhead}.
%$D(s)$: the maximum distance between centralized controller and CR users.
%
Figure~\ref{control_msg} shows the number of transmissions of SOC, the upper bound of the number of transmissions for ROSS, and the analytical number of transmissions of the centralised scheme. 
%Note that the curve for the centralized scheme is from theoretical analysis, which is $h+m+N$ as discussed beforehand.
%Note the overhead involved to construct the backbone (clusters) is not included.

%\begin{table}[hc]
%\centering
%\caption{Singalling overhead. Notations: $n$-number of CR nodes in CRN, $h$-number of cluster heads, $m$-number of debatable nodes, $c$-number of demanding clusters.}\label{tab_overhead}
%\begin{tabular}{|p{2.2 cm}|p{1.5 cm}|p{3.7 cm}|}
%\hline
% Scheme 					&   Number of transmissions 					& Content of message \\ \hline
% ROSS-DGA, ROSS-$\delta$-DGA 		&   $h+2*m^2c$ (upper bound)
%				& $ID_{H_C}$ and $K_C$ for $h+m^2c$ times, notification to join one cluster for $m^2c$ times					\\ \hline
% ROSS-DFA, ROSS-$\delta$-DFA 		&   $h+ 2m$	 (upper bound) 						& $ID_{H_C}$ and $K_C$ for $h+m$ times, notification to join one cluster for $m$ times	 					\\ \hline
% SOC 						&   $3*n$					& $\{K_i\}, i\in M\subseteq \texttt{Nb}_i$						\\ \hline
% Centralized				&	$n$						& $\{C\}$         	\\ 
% \hline
%\end{tabular}
%\end{table}

\begin{center}
\begin{table*}[!htb]
\caption{Signalling overhead}\label{tab_overhead}
{\small
\hfill{}
\begin{tabular}{|L{2.2 cm}|C{2.8 cm}|C{3.15 cm}|C{6.6 cm}|}
\hline
 Scheme 				&Message Complexity 	&   Quantitative number of messages 		& Content of message (size of message) 									\\ \hline
 ROSS-DGA, ROSS-$\delta$-DGA 	&$\mathcal{O}(N^3)$ (worst case)		&   $h+2*m^2d$ (upper bound)  				&   \multirow{2}{*}{\parbox{6.4cm}{Cluster head broadcasts cluster composition ($|C(i)|* |\mathcal{K}|$ bytes). Cluster members broadcast the manipulated new individual connectivity degree (1 byte)}}								\\ \cline{1-3}
 ROSS-DFA, ROSS-$\delta$-DFA 	&$\mathcal{O}(N)$ (worst case)		&   $h + 2m$	 (upper bound) 					& 	      												\\ \hline
 SOC 					&$\mathcal{O}(N)$		&   $3*N$									& $C(i)$ ($|C(i)|$ bytes) and $K(C(i))$ ($|\mathcal{P}|$ bytes), $i\in \mathcal{N}$						\\ \hline
 Centralized			&$\mathcal{O}(N)$			&	$h + m + N$ (upper bound)~\cite{Efficient_broadcasting_gathering_adhoc}		& $\{C\}$ ($|C_i|*N$ bytes)        					\\ \hline
\end{tabular}
}
\hfill{}
\end{table*}
\end{center}
% centralized: $\mathcal{O}(D(s)+log N)$ + $\mathcal{O}(1+D(s)+log N)$

\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.6\linewidth]{number_controlMsg.pdf}
  \caption{Number of control messages. Note the curves of ROSS-DGA and ROSS-DFA are the upper bounds}
  \label{control_msg}
\end{figure}

\subsection{Comparison among the Distributed Schemes}
\label{largeScaleCRN}
In this section we investigate the performances of distributed clustering schemes in CRN with different network scales and densities.
The transmission range of CR is $A/5$, PR's transmission range is $2A/5$.
The initial number of PU is 30.
%The number of CR is 100, 200 and 300, and the average number of neighbours of each CR is 9.5, 20, and 31.
The desired sizes adopted are listed in the Table~\ref{Simulation_para}, which is about 60\% of the average number of neighbours.

\begin{table}[ht]
\caption{}
\label{Simulation_para}
{\small
\hfill{}
\begin{tabular}{|L{3.7 cm}|C{1 cm}|C{1 cm}|C{1 cm}|}
\hline
Number of CRs			& 100 	&  200 					& 300 \\ \hline
Average num. of neighbours 	&9.5	&   20		& 31  \\ \hline
Desired size $\delta$ 	& 6	&   12 						& 20      \\ \hline
\end{tabular}
}
\hfill{}
\end{table}

\subsubsection{Number of CCs per Non-singleton Clusters}

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.7\linewidth]{ccc_large_scale_color.pdf}
  \caption{Number of common channels of non-singleton clusters.}
  \label{ccc_large_scale}
\end{figure}

Figure~\ref{ccc_large_scale} shows the average number of CCs of the non-singleton clusters.
We notice that SOC achieves the most CCs per non-singleton cluster, although the lead over the variants of ROSS shrinks significantly when $N$ increases.
%\ie when $N=300$, the number of CCs achieved by ROSS variants (except for ROSS-$\delta$-DFA) is almost the same as that resulted from SOC.
%This means SOC out performs in terms of the average number of CCs per non-singleton cluster when network is sparse.
%This is also observed in the evaluation in Section \ref{ccc_20} where $N=20$.
%When the network becomes denser, even this metric favours SOC as discussed in the beginning of Section~\ref{performance}, ROSS-DGA achieves even more CCs than SOC, and ROSS-DFA and ROSS-$\delta$-DGA increase the number of CCs visibly.

\subsubsection{Robustness of the clusters against newly added PUs}
We add extra 20 batches of PUs sequentially in the CRN, where each batch includes 10 PUs. 
%In this part of simulation, we investigate the robustness of clusters by increasing the PUs working on certain channels.
Figure~\ref{singleton_clusters_100} and \ref{singleton_clusters_200} show that when $N=100$ and 200, more unclustered CR nodes appear in the CRN where SOC is applied.
When the network becomes denser, ROSS-DGA/DFA generate slightly more unclustered CR nodes than SOC when new PUs are not many, but SOC's performance deteriorates quickly when the number of PUs becomes larger.
We only show the average values of the variants of ROSS as their confidence intervals overlap.
%
When applying ROSS with size control mechanism, significantly less unclustered CR nodes are generated.
Besides, the greedy mechanism moderately strengthens the robustness of the clusters.
%It can be seen that greedy algorithms result in slightly less singleton clusters than their counterparts.
%
%Figure ~\ref{singleton_clusters_300} depicts a denser CRN where $N=300$.
%SOC noticeably causes more singleton clusters than ROSS variants, except that ROSS-20-DFA results in more singleton clusters when PUs are few.
%The reason is ROSS-20-DFA conducts cluster membership clarification for only once, which causes large number of singleton clusters.
%ROSS-20-DGA increases the size of smaller clusters through debatable nodes' repeated updates thus drastically decreases the number of singleton clusters.
%\begin{figure}[!h]
%  \centering
%  \includegraphics[width=0.7\linewidth]{survival_rate_100_edge50.pdf}
%  \caption{Percentage of CRs which are not included in any clusters with the increasing number of primary users, $N=100$}
%  \label{singleton_clusters_100}
%\end{figure}
%  
%  \begin{figure}[!h]
%    \centering
%   \includegraphics[width=0.7\linewidth]{survival_rate_300_edge50.pdf}
%  \caption{Percentage of CRs which are not included in any clusters with the increasing number of primary users, $N=300$}
%  \label{singleton_clusters_300}
%\end{figure}

\begin{figure*}[t]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{survival_rate_100_edge50.pdf}\par\caption{100 CRs}\label{singleton_clusters_100}
    \includegraphics[width=\linewidth]{survival_rate_200_edge50.pdf}\par\caption{200 CRs}\label{singleton_clusters_200}
    \includegraphics[width=\linewidth]{survival_rate_300_edge50.pdf}\par\caption{300 CRs}\label{singleton_clusters_300}
\end{multicols}\caption{Percentage of CR nodes which are not included in any non-singleton clusters}
\label{unclustered_100_200_300}
\end{figure*}

%When the network is denser, the improvement on cluster sizes and robustness by the greedy search in the membership clarification phase is more obvious.

\begin{figure*}[t]
\begin{multicols}{3}
    \includegraphics[width=\linewidth]{cdf_clusterSize_100.pdf}\par\caption{100 CRs, 30 PUs in network}\label{cdf_clusterSize_100}
    \includegraphics[width=\linewidth]{cdf_clusterSize_200.pdf}\par\caption{200 CRs, 30 PUs in network}\label{cdf_clusterSize_200}
    \includegraphics[width=\linewidth]{cdf_clusterSize_300.pdf}\par\caption{300 CRs, 30 PUs in network}\label{cdf_clusterSize_300}
\end{multicols}
\caption{Cumulative distribution of CRs residing in clusters with different sizes}
\label{cdf_100_200_300}
\end{figure*}

\subsubsection{Cluster Size Control}
Figure~\ref{nClusters_largeNetwork} shows when the network density scales up, the number of formed clusters by ROSS increases by smaller margin, and that generated by SOC increases linearly.
This result coincides with the analysis in Section~\ref{ross_p2_cluster_pruning}.
%When the network becomes denser, more clusters are generated by SOC compared with ROSS variants.
To better understand the distribution of the sizes of formed clusters, we depict the empirical cumulative distribution of CR nodes which reside in clusters with certain sizes in CRNs in Figures~\ref{cdf_clusterSize_100}~\ref{cdf_clusterSize_200}~\ref{cdf_clusterSize_300}.

%In this group of evaluation, the number of PRs is 30.

\begin{figure}[!h]
  \centering
   \includegraphics[width=0.7\linewidth]{nClusters_largeNetwork.pdf}
  \caption{The number of formed clusters.}
  %, there are $x=6$ when $N=100$, $x=12$ when $N=200$, $x=21$ when $N=300$, which is around $2/3$ of the number of average neighbours.
  \label{nClusters_largeNetwork}
\end{figure}

%\begin{figure}[ht]
%\begin{center}
%%\centering
%\subfigure[100 CRs, 30 PRs]{\label{result1:1}\includegraphics[width=0.48\linewidth]{cdf_clusterSize_100.pdf}}
%\subfigure[300 CRns, 30 PRns]{\label{result1:2}\includegraphics[width=0.48\linewidth]{cdf_clusterSize_300.pdf}}
%\end{center} 
%\caption[Cluster sizes]{Distribution of CRs in clusters with different sizes} %{\subref{a}, \subref{b}, \subref{c}, \subref{d}}
%\label{result1}
%\end{figure}

The sizes of clusters generated by ROSS-DGA and ROSS-DFA span a wider range than ROSS with size control feature.
Most of the generated clusters are smaller than the average number of neighbours, which is roughly equal with the 95\% percentile of the ROSS-DGA curve.
The 50\% percentile of the ROSS-DGA curve is roughly the desired size $\delta$.
%
%The desired cluster sizes are 5, 8 and 13, which are around 45\% of the average number of neighbours in respective CRNs. 
When the variants of ROSS with size control feature are applied, the sizes of the most generated clusters are smaller than $\delta$.
%This means the size control feature effectively restricts the size of clusters, which is able to prevent the clusters from ungovernable growing with the increase of network density.
%But the range where most of the CRs (80\% which exclude the CR nodes residing in small and big clusters) resides centres a value, which is roughly the half of the average number of neighbours.
As to the curves of SOC, the 95\% percentiles are 36, 30, and 40 in respective networks.

Overviewing Figure~\ref{cdf_100_200_300}, we conclude that the sizes of the clusters generated by ROSS are limited by the network density, the sizes of the clusters formed by ROSS with size control feature are restricted by the desired size.
In contrary, the clusters generated from SOC demonstrate strong divergence on cluster sizes.

\subsection{Insights from the Simulation}
The centralized clustering scheme can form the clusters which satisfy the requirement on cluster size strictly, and the clusters are robust against the newly added PUs, besides, it generates the smallest control overhead in the process of cluster formation.

As distributed schemes, ROSS outperform the comparison scheme SOC considerately on three metrics.
ROSS generates much less singleton clusters than SOC, and the resulted clusters are robuster than SOC in front of newly added PUs.
The signalling overhead involved in ROSS is about half of that needed for SOC, and the signalling messages are much shorter that the latter.
The sizes of the clusters which are generated by ROSS are restricted by the network density, which demonstrate smaller discrepancy than that of SOC.
%
Besides, ROSS achieves similar performance to the centralized scheme, \ie the cluster sizes obtained by ROSS with size control feature are limited by the desired size, and the cluster robustness is similar when applying the variants of ROSS and the centralized scheme respectively.
As to the variants of ROSS, the greedy mechanism in ROSS-DGA helps to improve the performance on cluster size and cluster robustness at the cost of mildly increased signalling overhead.

We also notice that  as a metric, the number of CCs per non-singleton cluster doesn't indicate the robustness of clusters as shown in Figure~\ref{singleton_clusters} and \ref{unclustered_100_200_300}, although it is adopted as the metric in the formation of clusters.

\section{Conclusion}
\label{conclusion}
In this paper we investigate the robust clustering problem in CRN extensively.
The proposed scheme results in a cluster structure in CRN, which has longer time expectancy against the primary users' activity.
Besides, the clusters generated have similar sizes compared with other distributed schemes.
%generate robust clusters against primary users' influence, so that more CR users can benefit from operative decision making provided by the clusters.
We provide the mathematical description of the problem and prove the NP hardness of it.
We propose both centralized and distributed schemes, where the congestion game model in game theory is used to design the distributed schemes.
%A Light weighted clustering scheme ROSS is also proposed.
Through simulation and theoretical analysis, we find that distributed schemes achieve similar performance with centralized optimization in terms of cluster robustness, signaling overhead and cluster sizes, and outperform the comparison distributed scheme on the above mentioned metrics.
%, which is important to form clusters which maintains unbroken to the greatest extent possible under primary users' activity.
%These schemes outperform other distributed clustering scheme in terms of both cluster survival ratio and control overhead.

%The clusters resulted from ROSS-DGA and its faster version ROSS-DFA are less vulnerable compared with other distributed clustering schemes, and demonstrates similar survival rate with centralized scheme under primary users' influence.
%An light weighted cluster size control mechanism is contained in both ROSS-DGA and ROSS-DFA, which is advantageous for cooperative sensing and network operation with clusters.
%Furthermore, considerable less control messages are generated when compared with other clustering schemes.

The shortage of ROSS scheme is it doesn't generate big clusters whose sizes exceed the cluster head's neighborhood.
This problem is attributed to fact that ROSS forms clusters on the basis of cluster head's neighborhood, and doesn't involve interaction with the nodes outside its neighborhood.
In the other way around, forming big cluster which extends a cluster head's neighborhood has limited applications, as multiple hop communication and coordination are required within such clusters.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
%\section{Algorithm~\ref{alg0}}
\section*{}
\begin{algorithm}              % enter the algorithm environment
\caption{ROSS phase I: cluster head determination and initial cluster formation for CR node $i$}          % give the algorithm a caption
\label{alg0} 
\DontPrintSemicolon
\SetAlgoLined
\KwIn{$d_j, g_j, j\in \text{Nb}_i\setminus \text{CHs}$. Empty sets $\tau_1,\tau_2$}
\KwResult{Returning 1 means $i$ is cluster head, then $d_j$ is set to 0, $j\in \text{Nb}_i\setminus \text{CHs}$. returning 0 means $i$ is not CH.}

\If{$\nexists j\in \text{Nb}_i\setminus \text{CHs}$, such that $d_i \geq d_j$}{
	return 1;
	}
\eIf{$\exists j\in \text{Nb}_i\setminus CHs$, such that $d_i > d_j$}{
	return 0;}{
	\If{$\nexists j\in \text{Nb}_i\setminus CHs$, such that $d_j == d_i$}{
	$\tau_1 \leftarrow j$
	}
}
\If{$\nexists j\in \tau_1$, such that $g_i \leq g_j$}{
	return 1;
	}
\eIf{$\exists j\in \tau_1$, such that $g_i < g_j$}{
	return 0;
	}
	{\If{$\nexists j\in \tau_1$, such that $g_j == g_i$}{
		$\tau_2\leftarrow j$
		}
	}
\If{$\texttt{ID}_i$ is smaller than any $\texttt{ID}_j$, $j\in \tau_2\setminus i$}{
	return 1;
	}
	{return 0;
	}
\end{algorithm}

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
%\section*{Algorithm~\ref{alg_size_control_available_CCC}}
\section*{}
\begin{algorithm}               % enter the algorithm environment
\caption{ROSS phase I: cluster head guarantees the availability of CC (start from line 1) / cluster size control (start from line 2)}          % give the algorithm a caption
\label{alg_size_control_available_CCC}
\DontPrintSemicolon
\SetAlgoLined
\KwIn{Cluster C, empty sets $\tau_1, \tau_2$}
\KwOut{Cluster C has at least one CC, or satisfies the requirement on cluster size}
%\tcc*[r]{When to guarantee available CCs, execute from line 1, when to control cluster size, execute from line 2}
\While {$K_C =\emptyset$} {
\While{$|C|> \delta$}{
	%calculate $\lambda = \min_{i\in C, i\neq H_C}(|K_{H_C}\cap K_i|)$;\\
	\eIf{$\exists$ only one $i\in C\setminus H_C$, $i = \argmin(|K_{H_C}\cap K_i|)$}{
			$C=C\setminus i$;
		}{
				$\exists$ multiple $i$ which satisfies $i = \argmin(|K_{H_C}\cap K_i|)$;\\ $\tau_1\leftarrow i$;		
		}

	\eIf{$\exists$ only one $i\in \tau_1$, $i = \argmax(|\cap_{j\in C\setminus i} K_j|-|\cap_{j\in C} K_j|)$}{
		$C=C\setminus i$;
		}{
			%$\exists$ multiple $i$ which satisfies $i = \argmax(|\cap_{j\in C\setminus i} K_j|-|\cap_{j\in C} K_j|)$;\\
			$C=C\setminus i$, where $i = \argmin_{i\in \tau_1} \texttt{ID}_i $
			%$\texttt{ID}_i$ is smaller than any $\texttt{ID}_j$, $j\in \tau_2\setminus i$;
	}
}
}
\end{algorithm}

\section*{}
\begin{algorithm}               % enter the algorithm environment
\caption{Debatable node $i$ decides its affiliation in phase II of ROSS}
%, chooses one claiming cluster to stay and leaves all the other claiming clusters}          % give the algorithm a caption,  cluster to settle
\label{alg4}
\DontPrintSemicolon
\SetAlgoLined
\KwIn{all claiming clusters $C\in S_i$}
\KwOut{one cluster $C\in S_i$, node $i$ notifies all its claiming clusters in $S_i$ about its affiliation decision.
}
\While{$i$ has not chosen the cluster, or $i$ has joined cluster $\tilde{C}$, but $\exists C'\in S_i, C'\neq \tilde{C}$, which has $|K(C'\setminus i)|-|K(C')|<|K(C\setminus i)|-|K(C)|$}{
	\eIf{$\exists$ only one $C\in S_i$, $C = \argmin(|K(C\setminus i)| - |K(C)|)$}{
			return $C$;
		}{
				$\exists$ multiple $C\in S_i$ which satisfies $C = \argmin(|K(C\setminus i)| - |K(C)|)$;\\ 
				$\tau_1\leftarrow C$;		
		}
	\eIf{$\exists$ only one $C\in \tau_1$, $C = \argmax(K_{h_C}\cap K_i)$}{
			return $C$;
		}{
				$\exists$ multiple $C\in S_i$ which satisfies $C = \argmax(K_{h_C}\cap K_i)$;\\ 
				$\tau_2\leftarrow C$;		
		}
	\eIf{$\exists$ only one $C\in \tau_2$, $C = \argmin|C|)$}{
			return $C$;
		}{
				return $\argmin_{C\in \tau_2}h_C$;\\
		}
		}
\end{algorithm}

% use section* for acknowledgment

\section*{Proof of Theorem~\ref{clustering:theorem}}
\label{proof_clustering:theorem}
\begin{proof}
%Note that the formed cluster can be a singleton cluster, \ie cluster size is 1.
We consider a CRN which can be represented as a connected graph.
% (COMMENT: Why is it necessarily connected? No special case where the graph has two components? ANSWER(Di): It is possible for the graph to have two components. My point it, when theorem works in one component, it works in the complete graph.)
To simplify the discussion, we assume the secondary users have unique individual connectivity degrees. 
Each user has an identical ID and a neighborhood connectivity degree.
This assumption is fair as the neighborhood connectivity degrees and node ID are used to break ties in Algorithm~\ref{alg0}, when the individual connectivity degrees are unique, it is not necessary to use the former two metrics. 
%(COMMENT: What do you mean by breaking ties?? I don't see why your argument holds. ANSWER(DI): according to Algorithm 1, to decide whether node $i$ is a cluster head, the comparision on $i$'s certain degrees with its neighbors needs to be made. It is possible that $i$ has the same degree with some neighbors, then we need ID to decide whether $i$ is cluster head.)

%we assume every node has at least one neighbour.
For the sake of contradiction, let us assume there exist some secondary user $\alpha$ which is not included into any cluster.
Then there is at least one node $\beta\in \text{Nb}_\alpha$ such that $d_{\alpha} > d_{\beta}$. 
According to Algorithm~\ref{alg0}, $\delta$ is not included in any clusters, because otherwise $d_{\beta} = M$, a large positive integer.
Now, we distinguish between two cases: 
%Otherwise, node $\alpha$ is eligible to form a cluster. COMMENT: unnecessary sentence!
If $\beta$ becomes cluster head, node $\alpha$ is included, the assumption is not true.
If $\beta$ is not a cluster head, then $\beta$ is not in any cluster, we can repeat the previous analysis made on node $\alpha$, and deduce that node $\beta$ has at least one neighbouring node $\gamma$ with $d_{\gamma} < d_{\beta}$.
% (COMMENT: $\beta$ not being a cluster head and $\beta$ not belonging to any cluster is not the same statement!!! ANSWER(Di): I add some more words)
Till now, when there is no cluster head identified, the unclustered nodes, \ie $\alpha$, $\beta$ form a linked list, where their connectivity degrees monotonically decrease.
But this list will not continue to grow, because the minimum individual connectivity degree is zero, and the length of this list is upper bounded by the total number of nodes in the CRN.
An example of the formed node series is shown as Figure~\ref{lemma1}.

\begin{figure}[ht!]
  \centering
\includegraphics[width=0.6\linewidth]{lemma1.pdf}
% \includegraphics{figure1.pdf}
	\caption{The node series discussed in the proof of Theorem~\ref{clustering:theorem}, the deduction begins from node $\alpha$}
	\label{lemma1}
\end{figure}

In this example, node $\omega$ is at the tail of the list.
As $\omega$ does not have neighboring nodes with lower individual connectivity degree, $\omega$ becomes a cluster head.
Then $\omega$ incorporates all its one-hop neighbours (here we assume that every newly formed cluster has common channels), including the nodes which precede $\omega$ in the list.
The nodes which join a cluster set their individual connection degrees to $M$, which enables the node immediately precede in the list to become a cluster head.
In this way, cluster heads are generated from the tail of list to the head of the list, and all the nodes in the list are in at least one cluster, which contradicts the assumption that $\alpha$ is not included in any cluster.

If we see a secondary user \textit{becoming a cluster head}, or \textit{becoming a cluster member} as one step, as the length of the list of secondary users is not larger than $N$, there are $N$ steps for this scenario to form the initial clusters.

%we know that within at most $N$ steps, all the nodes belong to certain clusters. (COMMENT: No, we don't! Where was that shown?)
\end{proof}

%
\section*{Proof of Theorem~\ref{theorem1}}
\label{proof_theorem1}
\begin{proof}
We put the definition of weighted k-set packing problem here as it to be used in the analysis on the complexity of the centralised clustering problem.

\begin{mydef}
\label{def_kset_packing}
\textit{Weighted k-set packing.}

Given a finite set $\mathcal{G} = \{ g_1, \ldots , g_N \} \subsetneq \mathbb{N}$ of non-negative integers and a collection of sets $\mathcal{Q}=\{S_1,S_2,\cdots,S_m\}$ such that $S_i \subseteq \mathcal{G}$ for every $1 \leq i \leq N$.
Each set $S \in \mathcal{Q}$ is associated with a weight $\omega(S) \in \mathbb{R}$.
Further, we are given a threshold value $\lambda \in \mathbb{N}$.\\
The question is whether there exists a collection $\mathcal{S} \subseteq \mathcal{Q}$ such that $\mathcal{S}$ contains only pairwise disjoint sets, i.e., for all $S, S' \in \mathcal{S}$ with $S \neq S'$ it holds that $S \cap S' = \emptyset$, and the total weight of the sets in $\mathcal{S}$ is greater than $\lambda$, i.e., $\sum_{S \in \mathcal{S}} \omega(S) > \lambda$.
\end{mydef}

Weighted k-set packing is NP-hard when $k\geqslant 3$.~\cite{Computers_a_Intractability}

To prove that the centralized clustering problem is NP-hard, we reduce the NP-hard problem \textit{weighted k-set packing} to it to prove the former is at least as hard as the latter.
%, which means centralized clustering in CRN problem is as hard as weighted k-set packing to be solved.
%To complete the reduction, we need to conduct following two steps:
%\begin{itemize}
%\item Step 1: We show the existence a polynomial algorithm $\sigma$, by which any instance $\mathcal{S}$ of a weighted k-set packing can be transformed to a clustering solution $\sigma(\mathcal{S})$ which complies with Definition~\ref{def_centralized_clustering}.
%\item Step 2: Show that $\mathcal{S}$ is a \textit{yes} instance of weighted k-set packing if and only if $\sigma(\mathcal{S})$ is a \textit{yes} instance for CRN clustering problem.
%\end{itemize}
We show the existence of a polynomial-time algorithm $\sigma$ that transforms any instance $\mathcal{I}$ of a weighted $k$-set packing into an instance $\sigma(\mathcal{I})$) of centralized clustering problem.

%We continue to use the notation adopted in the problem definition in Section~\ref{centralized_opt}.
W.l.o.g. let set $\mathcal{G} = \{ 1, \ldots , N \}$.
%$\mathcal{Q}$ is a collection of sets $\{s_1, s_2,\cdots s_m\}$, each set is composed with certain elements in $\mathcal{G}$. REPEATING THE DEF??
%Assume $\mathcal{S}\subseteq \mathcal{Q}$ is one instance of weighted k-set packing, and the sets in $\mathcal{S}$ are disjoint.
%$\omega$ indicates the weight for each set $s$, $\omega:\mathcal{S}\rightarrow \mathbb{Z}^{+}$. COMMENT: again unnecessary. Already contained in the definition!
%As the number of common channel of clusters is always integer
%give new weight to them in following way
The polynomial algorithm $\sigma$ consists of three steps.

\begin{itemize}

\item In the first step, we transform the instance $\mathcal{I}$ to $\mathcal{I'}$ by adding dummy elements into each set in $\mathcal{I}$.
Formally, for each set $s_i\in \mathcal{I}$, we create $s_i' = s_i \times \{ 1,2\}$.
Then we change the dummy elements by adding $N$ on them.
 %for instance, given $s_i=\{1, 4, 6\}$, the dummy set $s_i'$ is $\{1,1',4,4',6,6' \}$.
%By doing this, we obtain the dummy sets and constitute the dummy instance $\mathcal{S}'$ based on $\mathcal{S}$.
The purpose of this transformation is to eliminate the set in $\mathcal{I}$, which has single element.
The weight of these sets remain unchanged, \ie $\omega(s_i') = \omega(s_i)$.
%After this transmission, there is no set with only one element.
%This transformation requires $\sum_{s_i \in \mathcal{S}} |s_i|$ steps.

\item Mapping the elements in $\mathcal{I'}$ to CR nodes on a two-dimensional Euclidean plane, and mapping the sets in $\mathcal{I'}$ to clusters.
There is a bijection between the union of sets in $\mathcal{I'}$ to the CRN which consists of the CR nodes, in particular, each element one-to-one corresponds to the CR node whose node ID equals to the element.
%Given a collection $\mathcal{Q}$, on basis of which we construct a CRN. (I DON'T UNDERSTAND THAT. What do you mean?)
%We prepare $N$ CR nodes who are labelled from 1 to $N$, put them on a 2 dimension space, and deem a pair of them can communicate if they appear in one same set in $\mathcal{Q}$.
%We define a set of CR nodes $\mathcal{N} = \{ 1, \ldots , N \}$
We further regulate that, each set in $\mathcal{I'}$ has a corresponding cluster, \ie there is a bijection between a set in $\mathcal{I'}$ and a set of node IDs in a cluster, besides, the number of CCs is equal to the weight of the corresponding set.

%%There is one situation deserving extra explanation in the second transformation, we adopt one example to discuss it.
%Assuming two sets in $\mathcal{Q}$ are $s_1=\{1,2\}$ and $s_2=\{1,2,3\}$, then their weights are 3 and 5 respectively. (THAT'S TOO HAND-WAVY.)
%%Their dummy sets are $s_1'=\{1,1,2,2\}$ and $s_2'=\{1,1,2,2,3,3,4,4\}$ and their new weights are 3 and 5 as before.
%We find it is impossible to map the sets into clusters in the same time (SORRY, which sets?? The sets you mentioned before are already clusters, isn't it?), because the number of CCs of the cluster which bases on $s_1$ should be no less than the cluster which bases on $s_2$, as the latter has one extra node compared with the former cluster.
%%When $s_1'$ and $s_2'$ are mapped to CRN respectively, there is contradiction between them, because the number of common channels of CR node group $\{1,1,2,2\}$ is smaller than that of $\{1,1,2,2,3,3,4,4\}$.
%But as to any instance of the solution to the weighted k-set packing problem, this contradiction doesn't happen because the instance $\mathcal{S}$ contains only disjoint sets, thus at most only one set of $s_1$ and $s_2$ appears in $\mathcal{S}$.%, then we can safely delete the connections based on the deleted set from the CRN, and the contradiction is eliminated.
%(THIS PARAGRAPH IS TOO DIFFICULT FOR ME TO UNDERSTAND.)

\item In this step, we transform the instance $\mathcal{I'}$ to a clustering proposal for CRN.
%We prepare a second pool of CR nodes which are identical with the CR nodes prepared in step 1, i.e., identical IDs and channel availabilities on them, we call these CR nodes dummy nodes.
We adjust the locations of the CR nodes which are mapped from the dummy elements in the instance $\mathcal{I'}$, and let them beside the CR nodes whose ID is exact $N$ smaller.
% (THAT DOESN'T MAKE SENSE. THEN THE ID IS NOT UNIQUE ANYMORE.) in the CRN built in step 1, and there is connection between the CR node and its dummy node (the one CR node and its dummy node can be seen as two transceivers at one node). (YOU DON'T NEED TO GIVE PRACTICAL MOTIVATION IN A HARDNESS-PROOF.)
Because of the dummy nodes, the clustering solution which corresponds to $\mathcal{I'}$ doesn't have singleton cluster.
This transformation requires $2\cdot\sum_{s_i \in \mathcal{I'}} |s_i|$ steps.
%
Now we look back to the elements in $\mathcal{Q}$, which don't appear in $\mathcal{I'}$. 
We one-to-one map these elements to CR nodes, and arbitrarily put these CR nodes in the plane, and these CR nodes become single node clusters.
According to the definition of clustering problem in CRN, the number of CCs in these single node clusters is 0.
These singleton clusters and the clusters mapped from $\mathcal{I'}$ constitute a clustering proposal, and finding these singleton clusters requires at most $N$ steps.
%We deliberately locate them so that at least one CR node can communicate with the rest of them.
%map every element in $\mathcal{S}$ into one CR node, \ie each integer corresponds to one CR node, in particular, that integer becomes the CR node's ID.
%As $s'$ is dummy set which has duplicated elements, we also deploy the the same CR node twice.
%
An example is shown in Table~\ref{no_hard_proof_instance}.
\begin{table}[h!]
     \begin{center}
     \begin{tabular}{ p{3cm}  p{4.5cm} }
     \toprule
      $\mathcal{N}$  & $\{0, 1,2,3,4,5,6, 7, 8, 9\}$\\ 
    \cmidrule(r){1-1}\cmidrule(lr){2-2}     
      $\mathcal{Q}$  & $\{ \{1\}, \{1, 5\}, \{1,2,4\}, \{2,3\}, \{4\} \}$\\ 
    \cmidrule(r){1-1}\cmidrule(lr){2-2}     
      Instance for Weighted k-set packing  & $\{ \{1\}, \{2,3\}, \{4\} \}$\\ 
    \cmidrule(r){1-1}\cmidrule(lr){2-2}
      Instance with dummy elements
      & 
		$\{ \, \{1,11\},\{2,12, 3, 13\}, \{ 4,14\} \, \}$
      \\    \cmidrule(r){1-1}\cmidrule(lr){2-2}
	Instance for clustering solution (dashed circles are dummy nodes)
	  &
	  %\raisebox{-\totalheight}{\includegraphics[width=0.8\linewidth]{np_hard_proof_dummy.pdf}}      
	  \begin{minipage}{.3\textwidth}
      \includegraphics[width=0.6\linewidth]{np_hard_proof_dummy.pdf}
    \end{minipage}
	  \\       \bottomrule
      \end{tabular}
      \caption{}
      \label{no_hard_proof_instance}
      \end{center}
      \end{table}

%As to duplicated elements, we also map them into a CR node, thus there exist CR nodes with identical ID.
%As a result, these CR nodes constitute a collection of CR nodes, and they can not be called as a CRN yet as there are not communication links drawn between them.
%Connections in CRN under this context is decided by physical conditions, which means to have a communication link, the two CR nodes should have common channels and locate within each other's transmission range.
%We further assume that all CR nodes locate in a way that any two pair of nodes has potential to be connected if their IDs are in the same set in $\mathcal{S}$, the instance for weighted k-set packing.
%\item Mere isolated nodes don't constitute network, thus we add connections in CRN based on the sets in $\mathcal{S}'$ sequentially.
%For each set $s'\in \mathcal{S}'$, we add connection between two CR nodes if their IDs can be found in $s'$.
%No connection exists between two CR nodes if their IDs don't appear in the same set in $\mathcal{S}$.

%We complete the polynomial algorithm $\sigma$ for the the transformation so far.
%The number of common channels of cluster $f$ is non-increasing function of cluster size, while, the weight of set in weighted k-set packing problem doesn't have this property.
%In weighted k-set packing, the weight of a set with smaller size could be larger than a set with more elements.
%But this difference doesn't hinder the transformation because an instance of weighted k-set packing contains only disjoint sets in $\mathcal{G}$.

\end{itemize}

We have crossed the hurdle of finding one polynomial algorithm $\sigma$ which transforms an instance of weighted k-set packing to an instance of the clustering problem in CRN.
%Now we look into the step 2 in the reduction.
When $\mathcal{I}$ is not an instance for weighted k-set packing problem due to the existence of joint sets, the corresponding cluster strategy is not a solution for the centralized clustering problem, as there are overlapped clusters.

When there are only disjoint sets in an instance $\mathcal{I}$ for weighted k-set packing, the sum of weights is identical to the sum of the numbers of CCs in the CRN mapped from $\mathcal{I'}$, even $\mathcal{I}$ contains sets which only have one element.
Thus, when a instance $\mathcal{I}$ for k-set packing problem is true, \ie the sum of weights is greater than $\lambda$, then in the CRN which is mapped from $\mathcal{I'}$, the sum of the numbers of CCs of the clusters is greater than $\lambda$.
In the other way around, when an instance if false for weighted k-set packing problem,
%let's assume the maximum sum of weights of all instances is $\sum_{s_i\in \mathcal{S}}\omega(s_i)=\delta < \lambda$. 
%The dummy set of each $s_i\in \mathcal{S}$ is mapped to cluster of CR nodes.
%Definition of CRN clustering regulates that the number of common channels is 0 when the cluster has only one node.
%As to $|s_i|=1$, the mapped cluster has two nodes, with one of them is the dummy CR node.
%Then number of common channels is on longer 0 but equals to the weight of corresponding set $s_i$.
%%Meanwhile, the number of common channels of all the other single node cluster is 0.
the sum of the numbers of CCs of the clusters in the mapped CRN is smaller than $\lambda$, and there is no clustering solution satisfying the centralized clustering problem.

%When an instance $\mathcal{I}$ contains joint sets, there will be overlapped clusters among the mapped clusters, then neither is solution to their corresponding problem.

Thus, weighted k-set packing can be reduced to clustering problem in CRN, then the latter problem is of NP-hard.
\end{proof}

\section*{Acknowledgment}

The authors would like to thank xxxx

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}
\bibliographystyle{IEEEtran}
\bibliography{../backmatter/myrefs}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:
%IEEEbiographynophoto
\begin{IEEEbiography}{Di Li}
received BE and MS degrees in control engineering from Zhejiang University and Shaanxi University of Science and Technology respectively in China.
He worked with James Gross for his PhD in RWTH Aachen University since 2010.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiography}{Erwin Fang}
Biography text here.
Biography text here.
Biography text here.
Biography text here.
Biography text here.
Biography text here.

\end{IEEEbiography}
%\vspace{-50 mm}
% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiography}{James Gross}
Biography text here.
Biography text here.
Biography text here.
Biography text here.
Biography text here.
Biography text here.
Biography text here.

\end{IEEEbiography}

\end{document}